---
title: "compare_four_models_linear_regression"
author: "Renee Hui Xin Ng"
date: "2025-04-02"
output: html_document
---

```{r setup, include=FALSE}
libraries <- c("dplyr", "tidyverse", "DescTools", "ggplot2", "readxl", "lme4"
               , "ModelMetrics", "merTools", "lmtest", "emmeans", "haven", "lmerTest"
               , "metaforest", "rstatix", "labelled", "data.table", 'irr')

lapply(libraries, require , character.only = TRUE)
df<-read.csv("../data_processed/df_all_brainage_2025.csv") 
 
df  <- df %>% mutate(Age2 = agevisit**2, Age3 = agevisit**3, 
                     bpad_ridge = ridge_BrainAge - agevisit,
                     bpad_photon = age_prediction - agevisit,
                     bpad_BrainAgeR = ageprediction_BrainAgeR - agevisit,
                     bpad_cnn = cnn_BrainAge - agevisit)
 
```


### Brain-PAD Residuals from linear model with age, sex and group as predictors
```{r}
mod <- lm(bpad_photon ~  agevisit + gender + agevisit*gender ,data = df) 
summary(mod)
resid_bpad_photon <- resid(mod)

mod <- lm(bpad_ridge ~ agevisit + gender  ,data = df)
summary(mod)
resid_bpad_ridge <- resid(mod)
 
mod <- lm(bpad_BrainAgeR ~  agevisit + gender + agevisit*gender ,data = df)
summary(mod)
resid_bpad_BrainAgeR <- resid(mod) 
mod <- lm(bpad_cnn ~ agevisit + gender  ,data = df)
summary(mod)
resid_bpad_cnn <- resid(mod)
```



### Model accounting for interaction between Dx and Age, Dx and Sex, and Age and Sex, and three way interaction between Dx Age and Sex in ALL four models
```{r}
ridgemod <- lm(bpad_ridge ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender + dxgroupFK*gender*agevisit  , data = df) 
summary(ridgemod)$coefficients  %>%  kableExtra::kable(digit = 3)

photonmod <- lm(bpad_photon ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender  +dxgroupFK*gender*agevisit , data = df) 
summary(photonmod)$coefficients  %>% kableExtra::kable(digit = 3)

brainageRmodThreeWay <- lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender +dxgroupFK*gender*agevisit  , data = df) 
summary(brainageRmodThreeWay)$coefficients  %>% kableExtra::kable(digit = 3)

cnnmodThreeWay <- lm(bpad_cnn ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender  +dxgroupFK*gender*agevisit , data = df) 
summary(cnnmodThreeWay)$coefficients  %>% kableExtra::kable(digit = 3)
```

### Model Selection for CNN model
model selected: lm(bpad_cnn~ dxgroupFK + gender +   agevisit, data = df) 
```{r}
cnnmodNoThreeWay <- lm(bpad_cnn~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender   , data = df) 
summary(cnnmodNoThreeWay)$coefficients  %>% kableExtra::kable(digit = 3)
anova(cnnmodNoThreeWay, cnnmodThreeWay, test = "Chisq")

cnnmodNoDxAge<- lm(bpad_cnn~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + agevisit*gender   , data = df) 
summary(cnnmodNoDxAge)$coefficients  %>% kableExtra::kable(digit = 3)
anova(cnnmodNoDxAge, cnnmodNoThreeWay, test = "Chisq")

cnnmodDxSex<- lm(bpad_cnn~ dxgroupFK + gender +   agevisit + dxgroupFK*gender   , data = df) 
summary(cnnmodDxSex)$coefficients  %>% kableExtra::kable(digit = 3)
anova(cnnmodDxSex,  cnnmodNoDxAge, test = "Chisq")

cnnmod <- lm(bpad_cnn~ dxgroupFK + gender +   agevisit   , data = df) 
summary(cnnmod)$coefficients  %>% kableExtra::kable(digit = 3)
anova(cnnmodDxSex,  cnnmod, test = "Chisq")
# cnnmodNoDxAge
```


### PHOTON Model
Model selected: lm(bpad_photon ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender  +dxgroupFK*gender*agevisit , data = df) 

```{r}
photonNoThreeWay <-lm(bpad_photon ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender , data = df)
summary(photonNoThreeWay)$coefficients  %>% kableExtra::kable(digit = 3)
anova(photonNoThreeWay, photonmod, test = "Chisq")

photonNoDxAge<- lm(bpad_photon~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + agevisit*gender   , data = df) 
summary(photonNoDxAge)$coefficients  %>% kableExtra::kable(digit = 3)
anova(photonNoDxAge, photonNoThreeWay, test = "Chisq")
# brainageRmodNoDxAge
```

### Multimodal Ridge Model
Model selected: lm(bpad_ridge ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender , data = df)

```{r}
ridgeNoThreeWay <-lm(bpad_ridge ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender , data = df)
summary(ridgeNoThreeWay)$coefficients  %>% kableExtra::kable(digit = 3)
anova(ridgeNoThreeWay, ridgemod, test = "Chisq")

ridgeNoDxAge<- lm(bpad_ridge~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + agevisit*gender   , data = df) 
summary(ridgeNoDxAge)$coefficients  %>% kableExtra::kable(digit = 3)
anova(ridgeNoDxAge, ridgeNoThreeWay, test = "Chisq")


ridgeNoSexAge<- lm(bpad_ridge~ dxgroupFK + gender +   agevisit + dxgroupFK*gender   , data = df) 
summary(ridgeNoSexAge)$coefficients  %>% kableExtra::kable(digit = 3)
anova(ridgeNoDxAge, ridgeNoSexAge, test = "Chisq")

ridgeNoInteraction<- lm(bpad_ridge~ dxgroupFK + gender +   agevisit   , data = df) 
summary(ridgeNoInteraction)$coefficients  %>% kableExtra::kable(digit = 3)
anova(ridgeNoInteraction, ridgeNoSexAge, test = "Chisq")
```


### Model Selection for BrainAgeR
model selected: lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit   , data = df) 
````{r}
brainageRmodNoThreeWay <- lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender   , data = df) 
summary(brainageRmodNoThreeWay)$coefficients  %>% kableExtra::kable(digit = 3)
anova(brainageRmodThreeWay, brainageRmodNoThreeWay, test = "Chisq")

brainageRmodNoDxAge<- lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + agevisit*gender   , data = df) 
summary(brainageRmodNoDxAge)$coefficients  %>% kableExtra::kable(digit = 3)
anova(brainageRmodNoDxAge, brainageRmodNoThreeWay, test = "Chisq")

brainageRmodDxSex<- lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit + dxgroupFK*gender   , data = df) 
summary(brainageRmodDxSex)$coefficients  %>% kableExtra::kable(digit = 3)
anova(brainageRmodNoDxAge,  brainageRmodDxSex, test = "Chisq")

brainageRmod <- lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit   , data = df) 
summary(brainageRmod)$coefficients  %>% kableExtra::kable(digit = 3)
anova(brainageRmod,  brainageRmodDxSex, test = "Chisq")


```

## Is brain-PAD predictive of dxgroup?
```{r}
mod<-glm(dxgroupFK~ bpad_cnn   + gender + agevisit  , data = df , family = "binomial") 
summary(mod)

mod<- glm(dxgroupFK ~ bpad_photon*agevisit*gender  , data = df , family = "binomial") 
summary(mod) 

mod<- glm( dxgroupFK ~ bpad_BrainAgeR + gender + agevisit , data = df, family = "binomial") 
summary(mod)

mod<- glm( dxgroupFK ~ bpad_ridge* gender +  agevisit*gender+  agevisit*bpad_ridge , data = df, family = "binomial") 
summary(mod)
```

```{r}
## absolute errors
df <- df %>% mutate(ae_ridge = abs(ridge_BrainAge - agevisit),
                     ae_photon = abs(age_prediction - agevisit),
                     ae_BrainAgeR = abs(ageprediction_BrainAgeR - agevisit),
                     ae_cnn = abs(cnn_BrainAge - agevisit))  

# Calculate MAE for each model
mae_ridge <- df %>%
  summarise(mae_ridge = mean(ae_ridge, na.rm = TRUE))

mae_photon <- df %>%
  summarise(mae_photon = mean(ae_photon))

mae_BrainAgeR <- df %>%
  summarise(mae_BrainAgeR = mean(ae_BrainAgeR))

mae_cnn <- df %>%
  summarise(mae_cnn = mean(ae_cnn))

# Combine the results into a single dataframe
result_df <- bind_rows(mae_ridge, mae_photon, mae_BrainAgeR, mae_cnn)
```


```{r}
### absolute errors
AE_ICC <- df %>% mutate(ae_ridge = ridge_BrainAge - agevisit,
                     ae_photon = abs(age_prediction - agevisit),
                     ae_BrainAgeR = abs(ageprediction_BrainAgeR - agevisit),
                     ae_cnn =abs(cnn_BrainAge - agevisit),
                     age2 = agevisit*2) 

res <-AE_ICC %>% dplyr::select(ae_ridge, ae_BrainAgeR) %>%  icc(., model = "oneway", type = "agreement")
print('UCSD Multimodal & BrainageR')
print(round(res$value,3))

res <-AE_ICC %>% dplyr::select(ae_photon, ae_ridge) %>%  icc(., model = "oneway", type = "agreement")
print('UCSD Multimodal & PHOTON')
print(round(res$value,3))

res <-AE_ICC %>% dplyr::select(ae_photon, ae_BrainAgeR) %>%  icc(., model = "oneway", type = "agreement")   
print('PHOTON & BrainageR')
print(round(res$value,3))


res <-df %>% dplyr::select(ridge_BrainAge, ageprediction_BrainAgeR) %>% icc(., model = "oneway", type = "agreement")  
print('UCSD Multimodal & BrainageR')
print(round(res$value,3))

res <-df %>% dplyr::select(age_prediction, ridge_BrainAge) %>% icc(., model = "oneway", type = "agreement")  
print('UCSD Multimodal & PHOTON')
print(round(res$value,3))

res <-df %>% dplyr::select(age_prediction, ageprediction_BrainAgeR) %>% icc(., model = "oneway", type = "agreement")   
print('PHOTON & BrainageR')
print(round(res$value,3))

res <- df %>% dplyr::select(bpad_ridge, bpad_BrainAgeR) %>%  icc(., model = "oneway", type = "agreement") 
print('UCSD Multimodal & BrainageR')
print(round(res$value,3))

res <- df %>% dplyr::select(bpad_photon, bpad_ridge) %>%  icc(., model = "oneway", type = "agreement")   
print('UCSD Multimodal & PHOTON')
print(round(res$value,3))

res <- df %>% dplyr::select(bpad_photon, bpad_BrainAgeR) %>%  icc(., model = "oneway", type = "agreement")
print('PHOTON & BrainageR')
print(round(res$value,3))

```






### Correlation between corrected brain-PAD residuals and age for each model; seperate plots by group; seperate lines by sex
```{r  }
resids_df <- data.frame(df$agevisit, df$gender, df$dxgroupFK ,resid_bpad_photon,  resid_bpad_BrainAgeR, resid_bpad_cnn) 
splot_titles <- c(resid_bpad_ridge = 'Multimodal', resid_bpad_photon = 'PHOTON', resid_bpad_BrainAgeR =  'BrainageR', resid_bpad_cnn = 'DenseNet',`0` = 'HC',  `1` = 'BD')

resids_df$RowID <- seq_len(nrow(resids_df))
  
resid_bpad_ridge_df <- data.frame(RowID = seq_along(resid_bpad_ridge), resid_bpad_ridge)

resids_df <- merge(resids_df, resid_bpad_ridge_df, by = "RowID", all.x = TRUE)
resids_df


resids_df %>% dplyr::select(resid_bpad_ridge, resid_bpad_photon , resid_bpad_BrainAgeR ,resid_bpad_cnn, df.dxgroupFK, df.gender, df.agevisit) %>%  melt(id = c('df.agevisit', 'df.dxgroupFK', 'df.gender')) %>%
  ggplot(aes(x = df.agevisit, y = value , color = factor(df.gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Corrected Brain-PAD") +
  theme_minimal() +
  facet_grid(~ variable + df.dxgroupFK, label  = as_labeller(splot_titles))+ #s and add annotations
  geom_text(
    data = . %>% group_by(variable,df.dxgroupFK ) %>% summarise(correlation = cor(df.agevisit, value, use = "complete.obs", method = "pearson")),
    aes(x = Inf, y = -Inf, label = paste("r =", round(correlation, 2)) ),
    hjust = 1, vjust = -1,
    color = "black"
  )   + #s and add annotations
  geom_text(
    data = . %>% group_by(variable,df.dxgroupFK) %>% summarise(pval = as.numeric(cor.test(df.agevisit, value)[3])),
    aes(x = Inf, y = -Inf, label = paste("p =", round(pval, 2)) ),
    hjust = 1, vjust = 0,
    color = "black"
  )   
ggsave("figures/bpadresid_corr_age_bysexdx.jpeg",dpi = 500, width = 12, height = 3)
dev.off()
```

### Correlation between corrected brain-PAD residuals and age for each model; seperate plots by group
```{r }
# resids_df <- data.frame(df$agevisit,   df$dxgroupFK, resid_bpad_ridge,resid_bpad_photon,  resid_bpad_BrainAgeR, resid_bpad_cnn)
splot_titles <- c(resid_bpad_ridge = 'Multimodal', resid_bpad_photon = 'PHOTON', resid_bpad_BrainAgeR =  'BrainageR', resid_bpad_cnn = 'DenseNet', `0` = 'HC',  `1` = 'BD')

resids_df %>% dplyr::select(resid_bpad_ridge, resid_bpad_photon , resid_bpad_BrainAgeR ,resid_bpad_cnn, df.dxgroupFK, df.agevisit) %>%  melt(id = c('df.agevisit', 'df.dxgroupFK' )) %>%
  ggplot(aes(x = df.agevisit, y = value )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Corrected Brain-PAD")  +
    scale_color_manual(values=c('black'))+
  theme_minimal() +
  facet_grid(~ variable + df.dxgroupFK, labeller = as_labeller(splot_titles))    + 
geom_text(data = . %>% group_by(variable, df.dxgroupFK) %>% summarise(correlation = cor(df.agevisit, value, use = "complete.obs", method = "pearson")) %>% mutate(correlation = round(correlation, 2)), aes(x = Inf, y = -Inf, label = paste("r =", correlation)), hjust = 1, vjust = -1,  color = "black") + 
geom_text(data = . %>% group_by(variable, df.dxgroupFK) %>% summarise(pval = as.numeric(cor.test(df.agevisit, value)[3]))%>% mutate(pval = round(pval, 3)), aes(x = Inf, y = -Inf, label =  paste('p = ',round(pval, 2)),hjust = 1, vjust = 0, color = "black" )) +
  theme(legend.position="none")

ggsave("figures/bpadresid_corr_age_bydx.jpeg",dpi = 500, width = 12, height = 3)
dev.off()

# Significance of correlations as indicated by p-val
# Only Multimodal's correlation between brain-PAD residuals and age is significant
resids_df %>% dplyr::select(resid_bpad_ridge, resid_bpad_photon , resid_bpad_BrainAgeR ,resid_bpad_cnn, df.dxgroupFK, df.agevisit) %>%  melt(id = c('df.agevisit', 'df.dxgroupFK' )) %>%group_by(variable, df.dxgroupFK) %>% summarise(pval = as.numeric(cor.test(df.agevisit, value)[3]))%>% mutate(pval = round(pval, 3)) 
```

### Correlation between corrected brain-PAD residuals and age for each model  
```{r fig.align="center"}

splot_titles <- c(resid_bpad_ridge = 'Multimodal', resid_bpad_photon = 'PHOTON', resid_bpad_BrainAgeR =  'BrainageR', resid_bpad_cnn = 'DenseNet')
# resids_df <- data.frame(df$agevisit,   df$dxgroupFK, resid_bpad_ridge, resid_bpad_photon,  resid_bpad_BrainAgeR, resid_bpad_cnn)


resids_df %>% dplyr::select(resid_bpad_ridge, resid_bpad_photon , resid_bpad_BrainAgeR ,resid_bpad_cnn,   df.agevisit) %>%  melt(id = c('df.agevisit'  )) %>%
  ggplot(aes(x = df.agevisit, y = value )) +
  geom_point() +theme_minimal()+
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Corrected Brain-PAD") +    scale_color_manual(values=c('black'))+
  facet_grid(~ variable, labeller = as_labeller(splot_titles)) + #s and add annotations
  geom_text(
    data = . %>% group_by(variable) %>% summarise(correlation = cor(df.agevisit, value, use = "complete.obs", method = "pearson")),
    aes(x = Inf, y = -Inf, label = sprintf("r = %.2e", correlation)), hjust = 1, vjust = -2,  color = "black"
  ) + #s and add annotations
  geom_text(
    data = . %>% group_by(variable) %>% summarise(pval = as.numeric(cor.test(df.agevisit, value)[3]))%>% mutate(pval = round(pval, 2)),
    aes(x = Inf, y = -Inf, label =  paste('p = ',pval),hjust = 1, vjust = -1 ,color = "black" )) +
  theme(legend.position="none")

ggsave("figures/corrected_bpadresid_corr_age.jpeg",dpi = 500, width = 12, height = 3)
dev.off() 

```
 