---
title: "compare_four_models_linear_regression"
author: "Renee Hui Xin Ng"
date: "2025-04-02"
output: html_document
---

# What does this script do?
 
Linear Models:
Fit linear regression models to obtain residuals for each algorithm (ridge, photon, BrainAgeR, CNN) using age and gender, and age*gender between these variables as covariates.
This allows assessment of how well the model's residuals are explained by these predictors.
Residualized Brain-PADs are then used to predict diagnosis group. 
We fit the most complex model and then choose the best fitting model by using anova. 

Model Selection:
Selects the best fitting models for each brain age model (CNN, photon, ridge, and BrainAgeR) based on the comparison results. Similar to ENIGMA-BD analyses. 
 
Correlation Analysis:
Correlation between the corrected brain-PAD/residualized brain-PAD and chronological age for each model. 
Showing the relationship between these brain-PAD and age, separated by diagnostic group (HC vs. BD) and sex. 
Pearson correlation coefficients and p-values are calculated and annotated on the plots.

Inter-Rater Reliability (ICC):
The script computes inter-rater reliability (using the icc() function from the irr package) between different brain age models (e.g., ridge vs. BrainAgeR, photon vs. ridge, etc.) to assess the consistency of the predictions across models.

Binomial Logistic Regression:
The script uses logistic regression to assess whether the brain-PAD residuals (or raw predictions) can predict diagnostic group (dxgroupFK). This tests if the models are informative for distinguishing between groups (e.g., healthy controls vs. those with brain disease).
 

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)       # used with dplyr pipelines
library(ggplot2)
library(reshape2)    # for melt() – even though deprecated, you're still using it
library(readxl)      # if reading Excel files, although not shown in session
library(haven)       # for `read_sav()` when loading SPSS files
library(DescTools)   # for `icc()` function
library(data.table)  # used with `setDT()` and `melt()`
library(lme4)        # for `lmer()`
library(lmerTest)    # for p-values in `lmer`
library(lmtest)      # optional, used for model testing
library(irr)         # for inter-rater reliability (e.g. `icc()`, alternative to 
library(pROC)
 
df<-read.csv("../data_processed/df_all_brainage_2025.csv") 
df  <- df %>% mutate(Age2 = agevisit**2, Age3 = agevisit**3, 
                     bpad_ridge = ridge_BrainAge - agevisit,
                     bpad_photon = age_prediction - agevisit,
                     bpad_BrainAgeR = ageprediction_BrainAgeR - agevisit,
                     bpad_cnn = cnn_BrainAge - agevisit)
 
```


### Obtain Brain-PAD Residuals from linear model with covariates of age and sex

Residualize brain age by adjusting for age, sex, or other  covariates before assessing its predictive value for the diagnostic group.
Examine the residuals in brain age predictions as explained by age and gender alone.  
Models are specifically focused on how these two factors influence brain age prediction residuals, independent of diagnosis group.
```{r}
mod <- lm(bpad_photon ~  agevisit  ,data = df) 
summary(mod)
resid_bpad_photon <- resid(mod)

mod <- lm(bpad_ridge ~ agevisit    ,data = df)
summary(mod)
resid_bpad_ridge <- resid(mod)
 
mod1 <- lm(bpad_BrainAgeR ~ agevisit  , data = df)
mod2 <- lm(bpad_BrainAgeR ~ agevisit + gender + agevisit:gender, data = df)
anova(mod1, mod2)
resid_bpad_BrainAgeR <- resid(mod1)


#summary(mod1)

 
mod1 <- lm(bpad_cnn ~ agevisit , data = df)
mod2 <- lm(bpad_cnn ~ agevisit + gender + agevisit:gender, data = df)
anova(mod1, mod2)
resid_bpad_cnn <- resid(mod1) 
summary(mod1)
 

```

### Integrate Residualized Brain-PADs from Ridge Multimodal model with the Residualized Brain-PADs from other models
(66 data points because HC=22, BD=44) with the remaining Residualized Brain-PADs from other models
````{r}
resids_df <- data.frame(df$agevisit, df$gender, df$dxgroupFK ,resid_bpad_photon,  resid_bpad_BrainAgeR, resid_bpad_cnn) 
splot_titles <- c(resid_bpad_ridge = 'Multimodal', resid_bpad_photon = 'PHOTON', resid_bpad_BrainAgeR =  'BrainageR', resid_bpad_cnn = 'DenseNet',`0` = 'HC',  `1` = 'BD')

resids_df$RowID <- seq_len(nrow(resids_df))
  
resid_bpad_ridge_df <- data.frame(RowID = seq_along(resid_bpad_ridge), resid_bpad_ridge)

resids_df <- merge(resids_df, resid_bpad_ridge_df, by = "RowID", all.x = TRUE)
length(resids_df)
```

### AUC
(66 data points because HC=22, BD=44) with the remaining Residualized Brain-PADs from other models
````{r}
mod<- glm(df.dxgroupFK ~ resid_bpad_cnn  + resid_bpad_BrainAgeR +resid_bpad_photon , data = resids_df , family = "binomial") 
summary(mod) 

resids_df

# ROC for the full model
full_model <- glm(df.dxgroupFK ~ resid_bpad_cnn  + resid_bpad_BrainAgeR + resid_bpad_photon,
                  family = "binomial", data = resids_df)
full_model
# Predict probabilities for full model
full_probs <- predict(full_model, type = "response")
full_probs

# Rewritten function using dplyr pipes
add_roc_curve <- function(var, color) {
  # Use dplyr to create model and get predictions
  model_data <- resids_df %>%
    dplyr::select(df.dxgroupFK, all_of(var)) %>%
    glm(formula = as.formula(paste("df.dxgroupFK ~", var)), 
        family = "binomial", 
        data = .)
  
  # Get predicted probabilities
  preds <- predict(model_data, type = "response")
  
  # Compute and plot ROC curve
  roc_obj <- roc(resids_df$df.dxgroupFK, preds)
  plot(roc_obj, add = TRUE, col = color, lwd = 2)
  
  return(round(auc(roc_obj), 3))
}


# Compute and plot ROC for full model
roc_full <- roc(resids_df$df.dxgroupFK, full_probs, ci = TRUE)
plot(roc_full, col = "black", lwd = 2, main = "ROC Curves for Residual Brain-PAD Predictors")

legend("bottomright", legend = paste("Full Model AUC:", round(auc(roc_full), 3)), col = "black", lwd = 2)

colors <- c("blue", "red", "green" )
vars <- c("resid_bpad_cnn",   "resid_bpad_BrainAgeR", "resid_bpad_photon")

# Run the AUC plotting
aucs <- mapply(add_roc_curve, vars, colors)
# Add legend
legend("topleft", legend = paste(vars, "AUC:", aucs), col = colors, lwd = 2)
ggsave("output/figures/roc_curve_three_models_resid_bpad.jpeg",
  width = 8, height = 6, dpi = 300)
```

```{r}
# Fit full model using raw Brain-PADs

full_model_raw <- glm(dxgroupFK ~    bpad_photon+ bpad_BrainAgeR + bpad_cnn,
                      family = "binomial", data = df)

summary(full_model_raw)
full_model_raw <- glm(dxgroupFK ~    bpad_cnn *agevisit* gender ,
                      family = "binomial", data = df)
summary(full_model_raw)
# Predict probabilities
full_probs_raw <- predict(full_model_raw, type = "response")
 # ROC for the full model
 roc_full_raw <- roc(df$dxgroupFK, full_probs_raw, ci = TRUE)
plot(roc_full_raw, col = "black", lwd = 2, main = "ROC Curves for across Brain-PAD Predictors")
legend("bottomright", legend = paste("Full Model AUC:", round(auc(roc_full_raw), 3)), col = "black", lwd = 2)

# Function using dplyr to plot ROC for each raw Brain-PAD
add_raw_roc_curve <- function(var, color) {
  model_data <- df %>%
    dplyr::select(dxgroupFK, all_of(var)) %>%
    glm(formula = as.formula(paste("dxgroupFK ~", var)),
        family = "binomial", data = .)
  
  preds <- predict(model_data, type = "response")
  roc_obj <- roc(df$dxgroupFK, preds)
  plot(roc_obj, add = TRUE, col = color, lwd = 2)
  return(round(auc(roc_obj), 3))
}

# Define variables and colors
colors <- c("blue", "red", "green")
vars <- c("bpad_cnn", "bpad_BrainAgeR", "bpad_photon")

# Plot individual ROC curves and get AUCs
aucs <- mapply(add_raw_roc_curve, vars, colors)

# Add legend
legend("topleft", legend = paste( c("DenseNet", "BrainageR", "PHOTON"), "AUC:", aucs), col = colors, lwd = 2)
ggsave("output/figures/roc_curve_three_models_bpad.jpeg",
  width = 8, height = 6, dpi = 300)
```


```{r}
# Fit full model using raw Brain-PADs

subset_hc_df <- df %>% filter(!is.na(bpad_ridge))
full_model_raw <- glm(dxgroupFK ~    bpad_photon + bpad_BrainAgeR + bpad_cnn + bpad_ridge,
                      family = "binomial", data = subset_hc_df)
summary(full_model_raw)
# Predict probabilities
full_probs_raw <- predict(full_model_raw, type = "response")
 # ROC for the full model
 roc_full_raw <- roc(subset_hc_df$dxgroupFK, full_probs_raw, ci = TRUE)
plot(roc_full_raw, col = "black", lwd = 2, main = "ROC Curves for across Brain-PAD Predictors")
legend("bottomright", legend = paste("Full Model AUC:", round(auc(roc_full_raw), 3)), col = "black", lwd = 2)

# Function using dplyr to plot ROC for each raw Brain-PAD
add_raw_roc_curve <- function(var, color) {
  model_data <- df %>%
    dplyr::select(dxgroupFK, all_of(var)) %>%
    glm(formula = as.formula(paste("dxgroupFK ~", var)),
        family = "binomial", data = .)
  
  preds <- predict(model_data, type = "response")
  roc_obj <- roc(df$dxgroupFK, preds)
  plot(roc_obj, add = TRUE, col = color, lwd = 2)
  return(round(auc(roc_obj), 3))
}

# Define variables and colors
colors <- c("blue", "red", "green", 'purple')
vars <- c("bpad_cnn", "bpad_BrainAgeR", "bpad_photon")

# Plot individual ROC curves and get AUCs
aucs <- mapply(add_raw_roc_curve, vars, colors)

# Add legend
legend("topleft", legend = paste( c("DenseNet", "BrainageR", "PHOTON", 'Multimodal'), "AUC:", aucs), col = colors, lwd = 2)
ggsave("output/figures/roc_curve_four_models_bpad.jpeg",
  width = 8, height = 6, dpi = 300)
```



## Is brain-pad or residualized brain-pad predictive of diagnosis group?
Determine if the raw Brain-PAD vs. corrected Brain-PAD (with covaraites age and gender and interaction term) predict the diagnostic group. 
Less interpretable when we have three way interactions that are significant. 

Using brain-pad residuals isolates the unique, unexplained variation in brain age that is not due to these covariates, which helps focus on brain pad variations that are more specifically linked to the disease. 
```{r}
mod<- glm(dxgroupFK ~ bpad_photon*agevisit*gender  , data = df , family = "binomial") 
summary(mod) 
mod<- glm(dxgroupFK ~ resid_bpad_photon  , data = df , family = "binomial") # residualized models come from adjusting age and gender (and interaction if applicable)
summary(mod) 

mod<- glm( dxgroupFK ~ bpad_BrainAgeR + gender , data = df, family = "binomial") 
summary(mod) 
mod<- glm( dxgroupFK ~ resid_bpad_BrainAgeR   , data = df, family = "binomial") 
summary(mod)

mod<- glm( dxgroupFK ~ bpad_ridge + gender +  agevisit + agevisit*gender , data = df, family = "binomial") 
summary(mod)
mod<- glm(df.dxgroupFK ~  resid_bpad_ridge, data = resids_df, family = "binomial") # note that source of df is different here
summary(mod)

mod<-glm(dxgroupFK ~ bpad_cnn  + gender + agevisit , data = df , family = "binomial") 
summary(mod)
mod<-glm(dxgroupFK ~ resid_bpad_cnn  , data = df , family = "binomial") 
summary(mod)
 
```
# Predict Brain-PAD using dxgroup adjusting for covariates sex and age

### Model accounting for interaction between Dx and Age, Dx and Sex, and Age and Sex, and three way interaction between Dx Age and Sex in ALL four models
```{r}
ridgemod <- lm(bpad_ridge ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender + dxgroupFK*gender*agevisit  , data = df) 
summary(ridgemod)#$coefficients  

photonmod <- lm(bpad_photon ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender  +dxgroupFK*gender*agevisit , data = df) 
summary(photonmod)#$coefficients  

brainageRmodThreeWay <- lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender +dxgroupFK*gender*agevisit  , data = df) 
summary(brainageRmodThreeWay)#$coefficients   

cnnmodThreeWay <- lm(bpad_cnn ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender  +dxgroupFK*gender*agevisit , data = df) 
summary(cnnmodThreeWay)#$coefficients   
```




### PHOTON Model

Model selected: lm(bpad_photon ~ dxgroupFK + gender +   agevisit + dxgroupFK* gender + dxgroupFK * agevisit + agevisit* gender  + dxgroupFK* gender*agevisit , data = df) 
```{r}
photonNoThreeWay <-lm(bpad_photon ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender , data = df)
summary(photonNoThreeWay)#$coefficients  
anova(photonNoThreeWay, photonmod, test = "Chisq")



photonNoDxAge<- lm(bpad_photon~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + agevisit*gender   , data = df) 
summary(photonNoDxAge)$coefficients   
anova(photonNoDxAge, photonNoThreeWay, test = "Chisq")
# brainageRmodNoDxAge
```

### Multimodal Ridge Model
Model selected: bpad_ridge ~ dxgroupFK + gender + agevisit + dxgroupFK * gender +  agevisit * gender

```{r}
ridgeNoThreeWay <-lm(bpad_ridge ~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender , data = df)
summary(ridgeNoThreeWay)$coefficients   
anova(ridgeNoThreeWay, ridgemod, test = "Chisq")

ridgeNoDxAge<- lm(bpad_ridge~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + agevisit*gender   , data = df) 
summary(ridgeNoDxAge)$coefficients  
anova(ridgeNoThreeWay , ridgeNoDxAge, test = "Chisq")


ridgeNoSexAge<- lm(bpad_ridge~ dxgroupFK + gender +   agevisit + dxgroupFK*gender   , data = df) 
summary(ridgeNoSexAge)$coefficients  
anova(ridgeNoDxAge, ridgeNoSexAge, test = "Chisq")

ridgeNoInteraction<- lm(bpad_ridge~ dxgroupFK + gender +   agevisit   , data = df) 
summary(ridgeNoInteraction)$coefficients  
anova(ridgeNoInteraction, ridgeNoSexAge, test = "Chisq")
```


### Model Selection for BrainAgeR
model selected: lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit   , data = df) 
````{r}
brainageRmodNoThreeWay <- lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender   , data = df) 
summary(brainageRmodNoThreeWay)$coefficients  
anova(brainageRmodThreeWay, brainageRmodNoThreeWay, test = "Chisq")

brainageRmodNoDxAge<- lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + agevisit*gender   , data = df) 
summary(brainageRmodNoDxAge)$coefficients   
anova(brainageRmodNoDxAge, brainageRmodNoThreeWay, test = "Chisq")

brainageRmodDxSex<- lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit + dxgroupFK*gender   , data = df) 
summary(brainageRmodDxSex)$coefficients  
anova(brainageRmodNoDxAge,  brainageRmodDxSex, test = "Chisq")

brainageRmod <- lm(bpad_BrainAgeR~ dxgroupFK + gender +   agevisit   , data = df) 
summary(brainageRmod)$coefficients  
anova(brainageRmod,  brainageRmodDxSex, test = "Chisq")



brainageRmodNoAge <- lm(bpad_BrainAgeR~ dxgroupFK  +   gender   , data = df) 
summary(brainageRmodNoAge)$coefficients  
anova(brainageRmod,  brainageRmodNoAge, test = "Chisq")
````
### Model Selection for CNN model
model selected: lm(bpad_cnn~ dxgroupFK + gender +   agevisit, data = df) 
```{r}
cnnmodNoThreeWay <- lm(bpad_cnn~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + dxgroupFK*agevisit + agevisit*gender   , data = df) 
summary(cnnmodNoThreeWay)$coefficients 
anova(cnnmodNoThreeWay, cnnmodThreeWay, test = "Chisq")

cnnmodNoDxAge<- lm(bpad_cnn~ dxgroupFK + gender +   agevisit + dxgroupFK*gender + agevisit*gender   , data = df) 
summary(cnnmodNoDxAge)$coefficients    
anova(cnnmodNoDxAge, cnnmodNoThreeWay, test = "Chisq")

cnnmodDxSex<- lm(bpad_cnn~ dxgroupFK + gender +   agevisit + dxgroupFK*gender   , data = df) 
summary(cnnmodDxSex)$coefficients   
anova(cnnmodDxSex,  cnnmodNoDxAge, test = "Chisq")

cnnmod <- lm(bpad_cnn~ dxgroupFK + gender +   agevisit   , data = df) 
summary(cnnmod)$coefficients  
anova(cnnmodDxSex,  cnnmod, test = "Chisq")
# cnnmodNoDxAge
```

### Mean Absolute Error for all four models
```{r}
## absolute errors
df <- df %>% mutate( ae_ridge = abs(agevisit - ridge_BrainAge ),
                     ae_photon = abs(age_prediction - agevisit),
                     ae_BrainAgeR = abs(ageprediction_BrainAgeR - agevisit),
                    ae_cnn = abs(cnn_BrainAge - agevisit))  
df %>%  filter(dxgroupFK == 0)%>% dplyr::select(ridge_BrainAge, age_prediction , cnn_BrainAge ,ageprediction_BrainAgeR, agevisit)  %>% cor(.)


# # if we want just HC = 22
# df <- df %>%filter(!is.na(ridge_BrainAge))%>% mutate( ae_ridge = abs(agevisit - ridge_BrainAge ),
#                      ae_photon = abs(age_prediction - agevisit),
#                      ae_BrainAgeR = abs(ageprediction_BrainAgeR - agevisit),
#                      ae_cnn = abs(cnn_BrainAge - agevisit))
 
#Calculate MAE for each model
mae_ridge <- df    %>% filter(dxgroupFK == 0)%>%
 summarise(mae_ridge = mean(ae_ridge, na.rm = TRUE))

mae_photon <- df %>% filter(dxgroupFK == 0)%>%
  summarise(mae_photon = mean(ae_photon))

mae_BrainAgeR <- df %>% filter(dxgroupFK == 0)%>%
  summarise(mae_BrainAgeR = mean(ae_BrainAgeR))

mae_cnn <- df %>% filter(dxgroupFK == 0)%>%
  summarise(mae_cnn = mean(ae_cnn))

# Combine the results into a single dataframe
result_df <- bind_rows(mae_ridge, mae_photon, mae_BrainAgeR, mae_cnn)
result_df

df %>%  filter(dxgroupFK == 0)%>% dplyr::select(ridge_BrainAge, age_prediction , cnn_BrainAge ,ageprediction_BrainAgeR, agevisit)  %>% cor(.)
 
 
```
 

```{r}
 plot_icc_heatmap <- function(icc_df, ordered_vars, custom_title) {
  custom_labels <- c("PHOTON", "BrainageR", "DenseNet", "Multimodal")

  icc_df_clean <- icc_df %>%
    filter(!is.na(Freq)) %>%
    mutate(
      Var1 = factor(Var1, levels = ordered_vars, labels = custom_labels),
      Var2 = factor(Var2, levels = ordered_vars, labels = custom_labels) 
    )
  
  ggplot(icc_df_clean, aes(x = Var1, y = Var2, fill = Freq)) +
    geom_tile(color = "white") +
    scale_fill_gradient2(
      low = "red", mid = "white", high = "darkgreen",
      midpoint = 0,
      name = "ICC"
    ) +
    geom_text(aes(label = sprintf("%.2f", Freq)), color = "black", size = 4) +
    theme_classic() +
    labs(title = custom_title, x = NULL, y = NULL) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Reorder variable names to match the new desired order
ordered_vars <- c("age_prediction", "ageprediction_BrainAgeR", "cnn_BrainAge", "ridge_BrainAge")

icc_mat <- compute_icc_matrix(df, ordered_vars)
icc_df <- as.data.frame(as.table(icc_mat)) %>% filter(!is.na(Freq))
title <- 'ICC between Predicted Ages across algorithms'

plot_icc_heatmap(icc_df, ordered_vars, title)

ggsave("output/figures/predicted_ages_icc_heatmap.jpeg",
       width = 8, height = 6, dpi = 300)
dev.off()


```
 
# BD ICC heatmap
```{r}
icc_mat <- df %>% filter(dxgroupFK == 1) %>% compute_icc_matrix(., c("ridge_BrainAge", "ageprediction_BrainAgeR", "cnn_BrainAge", "age_prediction"))
bd_icc_df <- as.data.frame(as.table(icc_mat)) %>% filter(!is.na(Freq))
ordered_vars <- c("age_prediction", "ageprediction_BrainAgeR", "cnn_BrainAge", "ridge_BrainAge")


title <- 'ICC between Predicted Ages across algorithms in BD'
plot_icc_heatmap(bd_icc_df, ordered_vars, title)
ggsave("output/figures/bd_predicted_ages_icc_heatmap.jpeg",
  width = 8, height = 6, dpi = 300)
dev.off()
```

# HC ICC heatmap
```{r}
icc_mat <- df %>% filter(dxgroupFK == 0) %>% compute_icc_matrix(., c("ridge_BrainAge", "ageprediction_BrainAgeR", "cnn_BrainAge", "age_prediction"))
hc_icc_df <- as.data.frame(as.table(icc_mat)) %>% filter(!is.na(Freq))
ordered_vars <- c("age_prediction",  "ageprediction_BrainAgeR", "cnn_BrainAge", "ridge_BrainAge")

title <- 'ICC between Predicted Ages across algorithms in HC'
plot_icc_heatmap(hc_icc_df, ordered_vars, title)
ggsave("output/figures/hc_predicted_ages_icc_heatmap.jpeg",
  width = 8, height = 6, dpi = 300)
dev.off()
```






### Correlation between corrected brain-PAD residuals and age for each model; facet plots by group; seperate lines by sex
```{r  }
resids_df <- data.frame(df$agevisit, df$gender, df$dxgroupFK ,resid_bpad_photon,  resid_bpad_BrainAgeR, resid_bpad_cnn) 
splot_titles <- c(resid_bpad_photon = 'PHOTON', resid_bpad_BrainAgeR =  'BrainageR', resid_bpad_cnn = 'DenseNet', resid_bpad_ridge = 'Multimodal', `0` = 'HC',  `1` = 'BD')

resids_df$RowID <- seq_len(nrow(resids_df))
resid_bpad_ridge_df <- data.frame(RowID = seq_along(resid_bpad_ridge), resid_bpad_ridge)

resids_df <- merge(resids_df, resid_bpad_ridge_df, by = "RowID", all.x = TRUE)

resids_df %>% dplyr::select(resid_bpad_ridge, resid_bpad_photon, resid_bpad_BrainAgeR ,resid_bpad_cnn, df.dxgroupFK, df.gender, df.agevisit) %>%  melt(id = c('df.agevisit', 'df.dxgroupFK', 'df.gender')) %>%
  mutate(
    df.gender = factor(df.gender, levels = c(0, 1), labels = c("Female", "Male")),
  #  dxgroupFK = factor(df.dxgroupFK, levels = c(0, 1), labels = c("HC", "BD")),
   # agevisit = df.agevisit
  )%>%
  ggplot(aes(x = df.agevisit, y = value , color = factor(df.gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Corrected Brain-PAD", color='Sex') +
  theme_minimal() +
  facet_grid(~ variable + df.dxgroupFK, label  = as_labeller(splot_titles) )+ #s and add annotations
  geom_text(
    data = . %>% group_by(variable,df.dxgroupFK ) %>% summarise(correlation = cor(df.agevisit, value, use = "complete.obs", method = "pearson")),
    aes(x = Inf, y = -Inf, label = paste("r =", round(correlation, 2)) ),
    hjust = 1, vjust = -1,
    color = "black"
  )   + #s and add annotations
  geom_text(
    data = . %>% group_by(variable,df.dxgroupFK) %>% summarise(pval = as.numeric(cor.test(df.agevisit, value)[3])),
    aes(x = Inf, y = -Inf, label = paste("p =", round(pval, 2)) ),
    hjust = 1, vjust = 0,
    color = "black"
  ) 
#ggsave("output/figures/bpadresid_corr_age_bysexdx.jpeg",dpi = 500, width = 12, height = 3)
dev.off()
```




### Correlation between corrected brain-PAD residuals and age for each model; facet plots by group (NO BRAINAGER)
```{r }
 
splot_titles <- c( resid_bpad_photon = 'PHOTON', resid_bpad_BrainAgeR =  'BrainageR', resid_bpad_cnn = 'DenseNet',resid_bpad_ridge = 'Multimodal', `0` = 'HC',  `1` = 'BD')

resids_df %>% dplyr::select(resid_bpad_photon,  resid_bpad_cnn,resid_bpad_ridge,  df.dxgroupFK, df.agevisit) %>%  melt(id = c('df.agevisit', 'df.dxgroupFK' )) %>%
  ggplot(aes(x = df.agevisit, y = value )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Corrected Brain-PAD")  +
    scale_color_manual(values=c('black'))+
  theme_minimal() +
  facet_grid(~ variable + df.dxgroupFK, labeller = as_labeller(splot_titles))    + 
geom_text(data = . %>% group_by(variable, df.dxgroupFK) %>% summarise(correlation = cor(df.agevisit, value, use = "complete.obs", method = "pearson")) %>% mutate(correlation = round(correlation, 2)), aes(x = Inf, y = -Inf, label = paste("r =", correlation)), hjust = 1, vjust = -1,  color = "black") + 
geom_text(data = . %>% group_by(variable, df.dxgroupFK) %>% summarise(pval = as.numeric(cor.test(df.agevisit, value)[3]))%>% mutate(pval = round(pval, 3)), aes(x = Inf, y = -Inf, label =  paste('p = ',round(pval, 2)),hjust = 1, vjust = 0, color = "black" )) +
  theme(legend.position="none")

ggsave("output/figures/no_brainager_bpadresid_corr_age_bydx.jpeg",dpi = 500, width = 12, height = 3)
dev.off()

# Significance of correlations as indicated by p-val
# Only Multimodal's correlation between brain-PAD residuals and age is significant
resids_df %>% dplyr::select(resid_bpad_ridge, resid_bpad_photon , resid_bpad_BrainAgeR ,resid_bpad_cnn, df.dxgroupFK, df.agevisit) %>%  melt(id = c('df.agevisit', 'df.dxgroupFK' )) %>%group_by(variable, df.dxgroupFK) %>% summarise(pval = as.numeric(cor.test(df.agevisit, value)[3]))%>% mutate(pval = round(pval, 3)) 
```

### Correlation between corrected brain-PAD residuals and age for each model
```{r fig.align="center"}

splot_titles <- c(resid_bpad_ridge = 'Multimodal', resid_bpad_photon = 'PHOTON', resid_bpad_BrainAgeR =  'BrainageR', resid_bpad_cnn = 'DenseNet')
# resids_df <- data.frame(df$agevisit,   df$dxgroupFK, resid_bpad_ridge, resid_bpad_photon,  resid_bpad_BrainAgeR, resid_bpad_cnn)


resids_df %>% dplyr::select( resid_bpad_photon  ,resid_bpad_cnn,resid_bpad_ridge,   df.agevisit) %>%  melt(id = c('df.agevisit')) %>%
  ggplot(aes(x = df.agevisit, y = value )) +
  geom_point() +theme_minimal()+
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Corrected Brain-PAD") +    scale_color_manual(values=c('black'))+
  facet_grid(~ variable, labeller = as_labeller(splot_titles)) + #s and add annotations
  geom_text(
    data = . %>% group_by(variable) %>% summarise(correlation = cor(df.agevisit, value, use = "complete.obs", method = "pearson")),
    aes(x = Inf, y = -Inf, label = paste("r =", round(correlation, 2))), hjust = 1, vjust = -2,  color = "black"
  ) + #s and add annotations
  geom_text(
    data = . %>% group_by(variable) %>% summarise(pval = as.numeric(cor.test(df.agevisit, value)[3]))%>% mutate(pval = round(pval, 2)),
    aes(x = Inf, y = -Inf, label =  paste('p = ',pval),hjust = 1, vjust = -1 ,color = "black" )) +
  theme(legend.position="none")


ggsave("output/figures/corrected_bpadresid_corr_age.jpeg",dpi = 500, width = 12, height = 3)
dev.off() 

```


```{r }
 
splot_titles <- c(resid_bpad_ridge = 'Multimodal', resid_bpad_photon = 'PHOTON', resid_bpad_BrainAgeR =  'BrainageR', resid_bpad_cnn = 'DenseNet', `0` = 'HC',  `1` = 'BD')

df %>% dplyr::select(ridge_BrainAge, age_prediction , ageprediction_BrainAgeR , cnn_BrainAge, df.dxgroupFK, df.agevisit) %>%  melt(id = c('df.agevisit', 'df.dxgroupFK' )) %>%
  ggplot(aes(x = df.agevisit, y = value )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Corrected Brain-PAD")  +
    scale_color_manual(values=c('black'))+
  theme_minimal() +
  facet_grid(~ variable + df.dxgroupFK, labeller = as_labeller(splot_titles))    + 
geom_text(data = . %>% group_by(variable, df.dxgroupFK) %>% summarise(correlation = cor(df.agevisit, value, use = "complete.obs", method = "pearson")) %>% mutate(correlation = round(correlation, 2)), aes(x = Inf, y = -Inf, label = paste("r =", correlation)), hjust = 1, vjust = -1,  color = "black") + 
geom_text(data = . %>% group_by(variable, df.dxgroupFK) %>% summarise(pval = as.numeric(cor.test(df.agevisit, value)[3]))%>% mutate(pval = round(pval, 3)), aes(x = Inf, y = -Inf, label =  paste('p = ',round(pval, 2)),hjust = 1, vjust = 0, color = "black" )) + theme(legend.position="none")

#ggsave("output/figures/bpadresid_corr_age_bydx.jpeg",dpi = 500, width = 12, height = 3)
#dev.off()
```

