---
title: "compare_four_models_cog_function"
author: "Hui Xin Ng"
date: "2023-07-15"
output: html_document
---

```{r setup, include=FALSE}
#renv::init()

pckgs <- c("dplyr", "tidyverse", "DescTools", "ggplot2", "readxl", "lme4"
               , "ModelMetrics", "merTools", "lmtest", "emmeans", "haven", "lmerTest",  "labelled", "data.table", 'outliers',    'Hmisc', 'irr', 'psych', 'ggtext', 'gridExtra')
pckgs %in% loadedNamespaces() # check if the packages are loaded

pacman::p_load(pckgs, character.only = TRUE)
pckgs %in% loadedNamespaces()

```

# HELPER FUNCTIONS
```{r}
# function to test whether data is normally distributed
test_normality <- function(data) {
  normality_test <- data %>%
  summarise(across(everything(), ~ shapiro.test(.)$p.value))
  return(normality_test)
}
## Helper functions
getPartialEtaSq<-function(m_fixed) {
  petasq <- round(as.data.frame(EtaSq(m_fixed))[2],3)
  return(petasq) }

# z transform data
z_transform <- function(data) {
  # Calculate column-wise means and standard deviations
  means <- colMeans(data)
  std_devs <- apply(data, 2, sd) 
  # Z-transform each column
  z_transformed_data <- scale(data, center = means, scale = std_devs) %>% as.data.frame()
  return(z_transformed_data)
}

# get linear model's coefficients
extract_coefficients_pvalues <- function(m, cog_score) {
  summary_age_mod <- summary(m) 
  coefficients <- summary_age_mod$coefficients[-1, "Estimate"] %>% round(2)
  p_values <- summary_age_mod$coefficients[-1, "Pr(>|t|)"]  
  
  # Replace p-values with significance levels or rounded values
  p_values_formatted <- ifelse(p_values < 0.001, "< 0.001",
                               ifelse(p_values < 0.01, "< 0.01",
                                      ifelse(p_values < 0.05, "< 0.05",
                                             round(p_values, 2)))) 
  result <- data.frame(coef_pval = paste(coefficients, '(', p_values_formatted,')'))
  rownames(result) <- c('sex', 'age', deparse(substitute(cog_score)), 'interaction' )
  result <- result %>% transpose() %>% `colnames<-`(c('sex', 'age', 'cog_score', 'interaction' ))
  return(result)
}

remove_outliers <- function(data, threshold = 3) {
  m <- apply(data, 2, mean)  # Calculate medians for each variable
  mad_values <- apply(data, 2, mad)  # Calculate median absolute deviations (MADs) for each variable
  
  # Calculate the outlier threshold as a multiple of MAD
  outlier_threshold <- threshold * mad_values
  
  # Identify outlier observations for each variable
  outliers <- apply(data, 2, function(x) abs(x - m) > outlier_threshold)
  
  # Remove outlier observations from the dataset
  cleaned_data <- data[!apply(outliers, 1, any), ]
  
  return(cleaned_data)
}
``` 
   
```{r cars}
path = file.path("../data_original/Master_BAI_080720_with_051517completers_060117imagingmaster_genotypes_brainages_ENIGMA_FS_MERGEFK_30122020.sav")

original_df = read_sav(path)
original_df%>% group_by(mri_scrn, dxgroup) %>% tally()


original_df <- original_df %>% dplyr::select(
agevisit,
dxgroupFK,
#trvsctrw, # DKEFS - Trails Visual Scanning Completion Times Raw  

trvsoerw, # DKEFS - Trails Visual Scanning Omission Errors Raw (Misses)
trvscerw, # DKEFS - Trails Visual Scanning Commission Errors Raw (False Alarm)

trnsctrw, # DKEFS - Trails Number Sequencing Completion Times Raw (TMT A)
trlnctrw, # DKEFS - Trails Letter-Number Completion Times Raw (TMT B)
trmsctrw, # DKEFS - Trails Motor Speed Completion Times Raw

cwicnrw, # DKEFS - Color Word Interference - Color Naming Raw / normally dist
cwiwrrw, # DKEFS - Color Word Interference - Word Reading Raw / normally dist

cwiirw, # DKEFS - Color Word Interference - Inhibition Raw
cwiisrw, # DKEFS - Color Word Interference - Inhibition/Switching Raw 

scid_a1, # SCID Current GAF
panspos, # PANS Positive Total
pansneg, # PANS Negative Total
pansgen, # PANS General Total
ymrstot, 
hamtot17,
hamtot28,

visit,
stdyptid) 
original_df <- setDT(original_df)[, lapply(.SD, na.omit), by = stdyptid]

original_df <- original_df %>% filter(visit == 0) %>% mutate(visit = 'Baseline')
 
 
#training_set <- original_df %>% dplyr::filter(!stdyptid %in% df$stdyptid, !stdyptid %in% c('BD087', 'BD223')) # only healthy individuals
original_df

```

 

```{r}
df<-read.csv("../data_processed/df_all_brainage_2025.csv") 
mod <- lm(dxgroup ~ agevisit, data = df)
summary(mod)
 
# this file is created from the Jupyter notebook to produce the same 44 BD and 37 HC participants for analyses  
df <- df %>% dplyr::select("stdyptid", 'gender','dxgroupFK' ,"visit", 'ridge_BrainAge',  'age_prediction', 'ageprediction_BrainAgeR' ,"cnn_BrainAge", "agevisit")  
 
df  <- df %>% mutate(bpad_ridge = ridge_BrainAge - agevisit,
                     bpad_photon = age_prediction - agevisit,
                     bpad_BrainAgeR = ageprediction_BrainAgeR - agevisit,
                     bpad_cnn = cnn_BrainAge - agevisit,
                     age2 = agevisit*2)
 

# remove NA values from each column in the data frame df while grouping the rows by the 'stdyptid' column
df <- setDT(df)[, lapply(.SD, na.omit), by = stdyptid] 

df$dxgroupFK <- haven::labelled(as.numeric(factor(df$dxgroupFK, levels = c("HC", "BD"))), labels = c('HC' = 0, 'BD' = 1))
# Drop 'visit' from original_df
 

 
original_df <- original_df %>%
  mutate(dxgroupFK = as.numeric(as_factor(dxgroupFK)))

df <- df %>%
  mutate(dxgroupFK = as.numeric(as_factor(dxgroupFK)))

# Now merge on stdyptid and dxgroupFK
 
merged_df <- merge(original_df, df, by = c("stdyptid", "agevisit"))
 
merged_df <- merged_df %>%
  rename(visit = visit.y,
         dxgroupFK = dxgroupFK.x)%>%
  dplyr::select(-visit.x, -dxgroupFK.y) %>%
  mutate(dxgroupFK = case_when(
    dxgroupFK == 1 ~ 0,
    dxgroupFK == 2 ~ 1,
    TRUE ~ dxgroupFK  # keep original value if not 1 or 2
  )) 
 
``` 


```{r} 
df <- merged_df 
df <- df %>%
  mutate(
    tmtb_a = trlnctrw - trnsctrw,
    stroop1 = cwiirw - (cwicnrw + cwiwrrw) / 2,
    stroop2 = cwiisrw - (cwicnrw + cwiwrrw) / 2,
    #gender = ifelse(gender == "Male", 1, 0),
    dxgroupFK = as_factor(dxgroupFK),
        gender = as_factor(gender) 

   # dxgroupFK = ifelse(dxgroupFK == "BD", 1, 0)
  )
df



ggplot(df, aes(x = agevisit, fill = factor(dxgroupFK))) +
  geom_density(alpha = 0.5) +
  labs(title = "Distribution of Ages Across BD and HC Groups",
       x = "Chronological Age",
       y = "Density",
       fill = "Group") + theme_classic()
```
# DATA CLEANING OF COGNITION VARIABLES
### Cognitive Performance Scores as predictor of brain-PAD 
### Cognitive scores distribution and summary stats of both BD and HC
```{r}
# Select column by index position
cog_test_scores  <- df %>% dplyr::select(trvsoerw, # DKEFS - Trails Visual Scanning Omission Errors Raw (Misses)
trvscerw, # DKEFS - Trails Visual Scanning Commission Errors Raw (False Alarm)

trnsctrw, # DKEFS - Trails Number Sequencing Completion Times Raw (TMT A)
trlnctrw, # DKEFS - Trails Letter-Number Completion Times Raw (TMT B)
trmsctrw, # DKEFS - Trails Motor Speed Completion Times Raw

cwicnrw, # DKEFS - Color Word Interference - Color Naming Raw / normally dist
cwiwrrw, # DKEFS - Color Word Interference - Word Reading Raw / normally dist

cwiirw, # DKEFS - Color Word Interference - Inhibition Raw
cwiisrw, # DKEFS - Color Word Interference - Inhibition/Switching Raw 
 stroop1, stroop2, tmtb_a ,stdyptid, dxgroupFK, gender)%>%  dplyr::select_if(~!all(is.na(.)))  # %>% filter(stroop1> 0, stroop2 >0, tmtb_a < 1500) #manual removal of outliers based on visualzation
# outliers are addressed in code beneath using mean absolute deviance 
# Reshape the dataframe from wide to long format
df_long <- reshape2::melt(cog_test_scores)

# Plot the distribution of cog scores using facet_wrap
ggplot(df_long, aes(x = value, fill = dxgroupFK)) +
  geom_histogram() +
  facet_wrap(~ variable, scales = "free") +
  theme_bw() 

summary_stats <- cog_test_scores %>% dplyr::select(-dxgroupFK, -stdyptid, -gender) %>%
  summarise(across(everything(), list(
    Mean = ~mean(.x, na.rm = TRUE),
    SD = ~sd(.x, na.rm = TRUE),
  #  Min = ~min(.x, na.rm = TRUE),
  #  Max = ~max(.x, na.rm = TRUE),
    Count = ~sum(!is.na(.x))
  ), .names = "{.col}_{.fn}"))

summary_stats %>% melt(.) %>% mutate(value = round(value, 2))
```
### Plot distribution of z-transformed non-normal cognitive scores
```{r}
cog_test_scores%>% dplyr::select(where(is.numeric)) %>% test_normality() > 0.05 # none of the scores are normal (when BD and HC are included) 
```
This section identifies and removes extreme outliers (more than 3 SD from the mean) from cognitive test scores using a standard deviation-based filter. It then:    
1. Visualizes the distributions of the cleaned variables.  
2. Computes summary statistics (mean and SD) for each variable.   
3. Tests for normality on numeric variables to determine suitability for parametric analyses.
```{r}
# there 5 subjects with scores > 3 std devs above the mean

# Remove extreme outliers (beyond ±3 standard deviations) and round remaining values
outliers_removed <- cog_test_scores %>%
  dplyr::select(-trvscerw, -trnsctrw) %>%  # Drop specific variables that are ranked or binary
  mutate_if(is.numeric, ~ ifelse(abs(. - mean(.)) / sd(.) <= 3, ., NA)) %>%
  dplyr::select_if(~!all(is.na(.))) %>%    # Drop columns that are all NA
  mutate_if(is.numeric, ~ round(., 3))

# Visualize cleaned distributions with histograms faceted by variable
outliers_removed %>%
  reshape2::melt() %>%
  ggplot(aes(x = value, fill = dxgroupFK)) +
  geom_histogram() +
  facet_grid(~ variable, scales = "free") +
  theme_bw()

# Compute mean and standard deviation for each numeric variable
summary_stats <- outliers_removed %>%
  dplyr::select(-dxgroupFK, -stdyptid, -gender) %>%
  summarise(across(everything(), list(
    Mean = ~mean(.x, na.rm = TRUE),
    SD = ~sd(.x, na.rm = TRUE)
  ), .names = "{.col}_{.fn}"))

# Reshape and round summary statistics for easier inspection
summary_stats %>%
  reshape2::melt() %>%
  mutate(value = round(value, 2))

# Optional: Check which numeric variables are approximately normally distributed (p > 0.05)
# Requires a custom function `test_normality()` applied across columns
outliers_removed %>%
  select_if(is.numeric) %>%
  test_normality > 0.05 

```


This script performs a PCA-based dimensionality reduction on cognitive task variables after removing outliers. It includes the following steps:

1. Adjusts skewed values by adding a constant to ensure all values are positive.
2. Removes variables that contain only NAs.
3. Applies a log10 transformation and standardizes numeric variables.
4. Flip signs for specific variables so that higher values reflect better cognitive performance.
5. Conducts Principal Component Analysis (PCA) and extracts the first principal component.
6. Merges the PCA scores with the original dataframe containing additional variables like age or diagnosis
```{r}
 

# Add a constant to shift all values of 'tmtb_a' to be positive
constant <- abs(min(outliers_removed$tmtb_a, na.rm = TRUE)) + 1
outliers_removed$tmtb_a <- outliers_removed$tmtb_a + constant

# Remove columns that are entirely NA
outliers_removed <- outliers_removed %>%
  select_if(function(x) !all(is.na(x)))

# Perform PCA on numeric variables after log transformation and scaling
pca_result <- outliers_removed %>%
  select_if(is.numeric) %>%
  log10() %>%
  mutate_if(is.numeric, scale) %>%
  mutate(
    cwicnrw = -cwicnrw,
    cwiwrrw = -cwiwrrw,
    cwiirw = -cwiirw,
    cwiisrw = -cwiisrw
  ) %>%
  principal(nfactors = 1)

# Create a dataframe that includes non-cognitive variables (e.g., age, diagnosis)
outliers_removed_df <- df %>% 
  filter(stdyptid %in% outliers_removed$stdyptid)

# Bind PCA scores to the full dataframe and flip direction so that higher = better performance
outliers_removed_df_pc <- cbind(as.data.frame(outliers_removed_df), pca_result$scores * -1)

# View the final dataframe with PCA scores
outliers_removed_df_pc
```

### cleaned dataset used for examining relationship between cognitive performance (PC1) and residualized brain-PADs from all four models
```{r}
write.csv(outliers_removed_df_pc, 'data_processed/outliers_removed_cognitive_test_scores.csv')

```

### Scree Plot
```{r}
# Extract eigenvalues from the PCA result
eigenvalues <- pca_result$values

# Calculate the proportion of variance explained by each principal component
variance_explained <- eigenvalues / sum(eigenvalues)

# Create a scree plot
plot(1:length(variance_explained), variance_explained, type = "b", 
     xlab = "Principal Component", ylab = "Proportion of Variance Explained",
     main = "Scree Plot")
variance_explained
```



# ANALYSIS
# Linear regression model with bpad as outcome variable
### Ridge Bpad
```{r}
# Filter out rows with missing bpad_ridge
filtered_df <- outliers_removed_df_pc %>%
  filter(complete.cases(bpad_ridge))

# Refit model using the filtered data
resid_mod <- lm(bpad_ridge ~ gender * agevisit, data = filtered_df)
summary(resid_mod)

# Get residuals from the model
filtered_df$resid_bpad_ridge <- resid(resid_mod)

# Regress PC1 on residualized bpad_photon
pca_cog_mod <- lm(PC1 ~ resid_bpad_ridge, data = filtered_df)
summary(pca_cog_mod)

# Get partial eta squared
getPartialEtaSq(pca_cog_mod)

m <- glm(dxgroupFK ~ agevisit, data = filtered_df, family = binomial)
summary(m)
```

### PHOTON Bpad
```{r}
resid_mod <- lm(bpad_photon ~   gender*agevisit, data = outliers_removed_df_pc)
resid_bpad_photon <- resid(resid_mod)
summary(resid_mod)
pca_cog_mod <- lm(PC1 ~  resid_bpad_photon  , data = outliers_removed_df_pc)
summary(pca_cog_mod)
getPartialEtaSq(pca_cog_mod) 
 
```


### BrainAgeR Bpad
```{r}
resid_mod <- lm(bpad_BrainAgeR ~   gender*agevisit, data = outliers_removed_df_pc)
resid_bpad_BrainAgeR <- resid(resid_mod)
grubbs.test(resid_bpad_BrainAgeR)
summary(resid_mod)
pca_cog_mod <- lm(PC1 ~  resid_bpad_BrainAgeR  , data = outliers_removed_df_pc)
summary(pca_cog_mod)
getPartialEtaSq(pca_cog_mod) 
 
```

```{r}
resid_mod <- lm(bpad_cnn ~   gender*agevisit, data = outliers_removed_df_pc)
resid_bpad_cnn <- resid(resid_mod)
grubbs.test(resid_bpad_cnn)
 
pca_cog_mod <- lm(PC1 ~ resid_bpad_cnn, data = outliers_removed_df_pc)
summary(pca_cog_mod) 
getPartialEtaSq(pca_cog_mod) 

 
``` 

### Four subplots showing the linear relationship between resid bpad and PC1

```{r}

# Calculate R-squared for the entire dataset
r2_values_BrainAgeR <- summary(lm(PC1 ~ resid_bpad_BrainAgeR, data = outliers_removed_df_pc))$r.squared

base_plot <- ggplot(outliers_removed_df_pc, aes(x = resid_bpad_BrainAgeR, y = PC1, color = factor(dxgroupFK))) +
  geom_point() + geom_smooth(aes(group = factor(dxgroupFK), color = factor(dxgroupFK)), method = "lm", formula = y ~ x, se = FALSE) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = 'darkorchid2') + # Add the purple linear regression line
  scale_color_manual(name = "Diagnosis Group", labels = c("HC", "BD"), values = c('darkblue', 'red')) +
  theme_bw() +
  theme(text = element_text(size = 11), legend.position = "bottom") +
  coord_cartesian(xlim = c(-20, 20), ylim = c(-3, 3))
r2_values_BrainAgeR <- summary(lm(PC1 ~ resid_bpad_BrainAgeR, data = outliers_removed_df_pc))$r.squared
# Annotate the plot with R-squared values for the first variable
plot1 <- base_plot +
  geom_richtext(aes(label = paste0("R<sup>2</sup> = ", format(round(r2_values_BrainAgeR, digits=3), nsmall = 2) ), x = Inf, y = -Inf, hjust = 1, vjust = -0.5), color = 'black') +
  labs(x = "Residual Brain-PAD", title = "BrainageR")

jpeg('figures/lm_PC1_resid_bpad_BrainAgeR.jpeg', width = 4000, height = 4000, units = "px", res = 500)
plot1
dev.off()
```


```{r}
base_plot <- ggplot(outliers_removed_df_pc, aes(x = resid_bpad_photon, y = PC1,  color = factor(dxgroupFK)))   +
  geom_point()  +
  geom_point() + geom_smooth(aes(group = factor(dxgroupFK), color = factor(dxgroupFK)), method = "lm", formula = y ~ x, se = FALSE) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = 'darkorchid2') + # Add the purple linear regression line
  scale_color_manual(name = "Diagnosis Group", labels = c("HC", "BD"), values = c('darkblue', 'red')) +
  theme_bw() +
  theme(text = element_text(size = 11), legend.position = "bottom") +
  coord_cartesian(xlim = c(-20, 20), ylim = c(-3, 3))
r2_values_photon <- summary(lm(PC1 ~ resid_bpad_photon, data = outliers_removed_df_pc))$r.squared

plot2 <- base_plot +
 geom_richtext(aes(label = paste0("R<sup>2</sup> = ",  format(round(r2_values_photon, digits=3), nsmall = 2) ), x = Inf, y = -Inf, hjust = 1, vjust = -0.5), color = 'black') +
  labs(x = "Residual Brain-PAD",title = "PHOTON-BA")



jpeg('figures/lm_PC1_resid_bpad_photon.jpeg', width = 4000, height = 4000, units = "px", res = 500)
plot2
dev.off()
```
 

```{r}

base_plot <- ggplot(outliers_removed_df_pc, aes(x = resid_bpad_cnn, y = PC1, color = factor(dxgroupFK)))   +
  geom_point() + geom_smooth(aes(group = factor(dxgroupFK), color = factor(dxgroupFK)), method = "lm", formula = y ~ x, se = FALSE) + geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = 'darkorchid2') + # Add the purple linear regression line
  scale_color_manual(name = "Diagnosis Group", labels = c("HC", "BD"), values = c('darkblue', 'red')) +
  theme_bw() +
  theme(text = element_text(size = 11), legend.position = "bottom") +
  coord_cartesian(xlim = c(-20, 20), ylim = c(-3, 3))


r2_values_cnn <- summary(lm(PC1 ~ resid_bpad_cnn, data = outliers_removed_df_pc))$r.squared



jpeg('figures/lm_PC1_resid_bpad_cnn.jpeg', width = 4000, height = 4000, units = "px", res = 500)
plot3 <- base_plot +  geom_richtext(aes(label = paste0("R<sup>2</sup> = ", format(round(r2_values_cnn, digits=3), nsmall = 2) ), x = Inf, y = -Inf, hjust = 1, vjust = -0.5), color = 'black') + labs(x = "Residual Brain-PAD",title = "MIDI")
plot3
dev.off() 

```

```{r}

base_plot <- ggplot(filtered_df, aes(x = resid_bpad_ridge, y = PC1, color = factor(dxgroupFK)))   +
  geom_point() + geom_smooth(aes(group = factor(dxgroupFK), color = factor(dxgroupFK)), method = "lm", formula = y ~ x, se = FALSE) + geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = 'darkorchid2') + # Add the purple linear regression line
  scale_color_manual(name = "Diagnosis Group", labels = c("HC", "BD"), values = c('darkblue', 'red')) +
  theme_bw() +
  theme(text = element_text(size = 11), legend.position = "bottom") +
  coord_cartesian(xlim = c(-20, 20), ylim = c(-3, 3))

r2_values_ridge <- summary(lm(PC1 ~ resid_bpad_ridge, data = filtered_df ))$r.squared
plot4 <- base_plot +
  geom_richtext(aes(label = paste0("R<sup>2</sup> = ", format(round(r2_values_ridge, digits=3), nsmall = 2) ), x = Inf, y = -Inf, hjust = 1, vjust = -0.5), color = 'black') +
  labs(x = "Residual Brain-PAD", title = 'UCSD Multimodal')

jpeg('figures/lm_PC1_resid_bpad_ridge.jpeg', width = 4000, height = 4000, units = "px", res = 500)
 
plot4
dev.off() 
```
 
```{r}
# Reshape the dataframe from wide to long format
df_long <- reshape2::melt(cog_test_scores)

# Open JPEG device
jpeg("figures/cog_performance_scores_distribution.jpeg", width = 1200, height = 800, quality = 100)

# Plot and save
ggplot(df_long, aes(x = value, fill = dxgroupFK)) +
  geom_histogram() +
  facet_wrap(~ variable, scales = "free") +
  theme_bw()

# Close the device
dev.off()

```

### Log transform cog scores
```{r}
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3577045/ PCA to run on DKEFS
# Ensure constant is computed properly
constant <- abs(min(outliers_removed$tmtb_a, na.rm = TRUE)) + 1

# Shift tmtb_a to be all positive
outliers_removed$tmtb_a <- outliers_removed$tmtb_a + constant

# Add a constant to the scores to make them positive
log_transformed <- outliers_removed %>%  mutate_if(., is.numeric, log10)
log_transformed %>%
  dplyr::select(where(is.numeric)) %>%
  dplyr::select(where(~ sum(!is.na(.)) >= 3)) %>%
  summarise(across(everything(), ~shapiro.test(.)$p.value))
log_transformed %>% reshape2::melt(.) %>%ggplot(., aes(x = value, fill=dxgroupFK)) +
  geom_histogram() +
  facet_wrap(~ variable, scales = "free") +
  theme_classic()

# Specify the columns to select from df2
selected_columns <- c("dxgroupFK", "gender", "agevisit", "stdyptid", 'bpad_photon', 'bpad_ridge', 'bpad_BrainAgeR', 'bpad_cnn', 'age_prediction', 'ageprediction_BrainAgeR', 'ridge_BrainAge', 'cnn_BrainAge')


```


