---
title: "compare_four_models_cog_function"
author: "Hui Xin Ng"
date: "2023-07-15"
output: html_document
---


```{r}


#renv::install(pkgs)
#renv::snapshot()
renv::status()
renv::restore()

# =========================
# Data import / pre processing
# =========================

# Read source data
path <- file.path(
  "..", "data_original",
  "Master_BAI_080720_with_051517completers_060117imagingmaster_genotypes_brainages_ENIGMA_FS_MERGEFK_30122020.sav"
)

library(dplyr)
original_df <- haven::read_sav(path)
original_df <- original_df %>%
  dplyr::mutate(
    BDI = ifelse(dxgroupFK == 0, 0, BDI)
  )

# Keep only variables used downstream
original_df <- original_df %>%
  dplyr::select(
    BDI,
    agevisit,
    dxgroupFK,

    # DKEFS Trails
    trvsoerw,   # visual scanning omission errors
    trvscerw,   # visual scanning commission errors
    trnsctrw,   # number sequencing completion time (TMT A)
    trlnctrw,   # letter-number sequencing completion time (TMT B)
    trmsctrw,   # motor speed completion time

    # DKEFS Color-Word Interference
    cwicnrw,    # color naming raw
    cwiwrrw,    # word reading raw
    cwiirw,     # inhibition raw
    cwiisrw,    # inhibition/switching raw

    # Clinical measures
    scid_a1,    # current GAF
    panspos,    # PANSS positive
    pansneg,    # PANSS negative
    pansgen,    # PANSS general
    ymrstot,
    hamtot17,
    hamtot28,
    HistoryPsychosis,

    visit,
    stdyptid
  )

s# Remove NAs within each participant, then keep baseline only
# Note: this preserves your original approach exactly
original_df <- setDT(original_df)[, lapply(.SD, na.omit), by = stdyptid] 
original_df <- original_df %>% filter(visit == 0) %>% mutate(visit = 'Baseline') 


```
 
### Load df_all_brainage_2025.csv to obtain the Brain Ages from all four algorithms
```{r}
# =========================
# Load brain-age data
# =========================

# This file was created from the Jupyter notebook to retain the same 44 BD and 37 HC participants used in the analyses
df <- read.csv("../data_processed/df_all_brainage_2025.csv")

# Keep only variables needed for merging and Brain-PAD calculation
df <- df %>%
  dplyr::select(
    stdyptid,
    gender,
    dxgroupFK,
    visit,
    ridge_BrainAge,
    age_prediction,
    ageprediction_BrainAgeR,
    cnn_BrainAge,
    agevisit
  )



# Create Brain-PAD variables and a doubled-age term
df <- df %>%
  dplyr::mutate(
    bpad_ridge     = ridge_BrainAge - agevisit,
    bpad_photon    = age_prediction - agevisit,
    bpad_BrainAgeR = ageprediction_BrainAgeR - agevisit,
    bpad_cnn       = cnn_BrainAge - agevisit,
    age2           = agevisit * 2
  )

# Remove missing values within each participant, this code was the same as above for original_df
df <- data.table::setDT(df)[, lapply(.SD, na.omit), by = stdyptid]

# =========================
# Harmonize diagnosis coding before merge
# =========================

# Convert diagnosis to numeric consistently in both datasets
original_df <- original_df %>%
  dplyr::mutate(dxgroupFK = as.numeric(haven::as_factor(dxgroupFK)))

df <- df %>%
  dplyr::mutate(dxgroupFK = as.numeric(haven::as_factor(dxgroupFK)))
```

### Merge clinical variables with brain age df
```{r}
# =========================
# Merge datasets
# =========================

# Merge on participant ID and age at visit
merged_df <- merge(original_df, df, by = c("stdyptid", "agevisit"))

# Keep the visit variable from df, the diagnosis variable from original_df,
# and recode diagnosis from 1/2 to 0/1 if needed
merged_df <- merged_df %>%
  dplyr::rename(
    visit = visit.y,
    dxgroupFK = dxgroupFK.x
  ) %>%
  dplyr::select(-visit.x, -dxgroupFK.y) %>%
  dplyr::mutate(
    dxgroupFK = dplyr::case_when(
      dxgroupFK == 1 ~ 0,
      dxgroupFK == 2 ~ 1,
      TRUE ~ dxgroupFK
    )
  )
```


### Check for group differences in terms of age and sex
```{r}
# 1. Plot age density by group
plot_age_density <- function(data, age_col = "agevisit", group_col = "dxgroupFK") {
  ggplot(data, aes_string(x = age_col, fill = paste0("factor(", group_col, ")"))) +
    geom_density(alpha = 0.5) +
    labs(
      title = "Distribution of Ages Across Groups",
      x = "Chronological Age",
      y = "Density",
      fill = "Group"
    ) +
    theme_classic()
}

# 2. Wilcoxon test for age by group
run_age_wilcox <- function(data, age_col = "agevisit", group_col = "dxgroupFK") {
  formula <- as.formula(paste(age_col, "~ as.factor(", group_col, ")", sep = ""))
  wilcox.test(formula, data = data)
}


# 5. Chi-square test for gender by group
run_gender_chisq <- function(data, group_col = "dxgroupFK", gender_col = "gender") {
  gender_table <- table(data[[group_col]], data[[gender_col]])
  chisq.test(gender_table)
}
```
```{r}
# =========================
# Simple age matching (no extra packages)
# =========================

# Split groups
bd <- merged_df %>% dplyr::filter(dxgroupFK == 1)   # BD
hc <- merged_df %>% dplyr::filter(dxgroupFK == 0)   # HC

original_df %>% dplyr::filter(dxgroupFK == 2)
original_df %>% dplyr::filter(dxgroupFK == 1)
# Sort by age (helps stability)
bd <- bd %>% arrange(agevisit)
hc <- hc %>% arrange(agevisit)

matched_pairs <- list()
used_hc <- rep(FALSE, nrow(hc))

for (i in seq_len(nrow(bd))) {
  age_bd <- bd$agevisit[i]
  
  # Compute distance to all unused HC
  diffs <- abs(hc$agevisit - age_bd)
  diffs[used_hc] <- Inf
  
  # Find closest HC
  j <- which.min(diffs)
  
  if (is.finite(diffs[j])) {
    matched_pairs[[length(matched_pairs) + 1]] <- rbind(bd[i, ], hc[j, ])
    used_hc[j] <- TRUE
  }
}

# Combine matched dataset
matched_df <- do.call(rbind, matched_pairs)

plot_age_density(matched_df)
run_gender_chisq(matched_df)
run_age_wilcox(matched_df)

for (df in list(matched_df, merged_df)) {
  print(plot_age_density(df))
  print(run_gender_chisq(df))
  print(run_age_wilcox(df))
}

```

# DATA CLEANING OF COGNITION VARIABLES
### Cognitive Performance Scores as predictor of brain-PAD 
### Cognitive scores distribution and summary stats of both BD and HC
```{r}
# Select column by index position
df
cog_test_scores  <- df %>% dplyr::select(trvsoerw, # DKEFS - Trails Visual Scanning Omission Errors Raw (Misses)
trvscerw, # DKEFS - Trails Visual Scanning Commission Errors Raw (False Alarm)

trnsctrw, # DKEFS - Trails Number Sequencing Completion Times Raw (TMT A)
trlnctrw, # DKEFS - Trails Letter-Number Completion Times Raw (TMT B)
trmsctrw, # DKEFS - Trails Motor Speed Completion Times Raw

cwicnrw, # DKEFS - Color Word Interference - Color Naming Raw / normally dist
cwiwrrw, # DKEFS - Color Word Interference - Word Reading Raw / normally dist

cwiirw, # DKEFS - Color Word Interference - Inhibition Raw 
cwiisrw, # DKEFS - Color Word Interference - Inhibition/Switching Raw 
 stroop1, stroop2, tmtb_a ,stdyptid, dxgroupFK, gender)%>%  dplyr::select_if(~!all(is.na(.)))  # %>% filter(stroop1> 0, stroop2 >0, tmtb_a < 1500) #manual removal of outliers based on visualzation
# outliers are addressed in code beneath using mean absolute deviance 
# Reshape the dataframe from wide to long format
df_long <- reshape2::melt(cog_test_scores)

# Plot the distribution of cog scores using facet_wrap
ggplot(df_long, aes(x = value, fill = dxgroupFK)) +
  geom_histogram() +
  facet_wrap(~ variable, scales = "free") +
  theme_bw() 

summary_stats <- cog_test_scores %>% dplyr::select( -stdyptid, -gender) %>% group_by(dxgroupFK) %>%
  summarise(across(everything(), list(
    Mean = ~mean(.x, na.rm = TRUE),
    SD = ~sd(.x, na.rm = TRUE)
  #  Min = ~min(.x, na.rm = TRUE),
  #  Max = ~max(.x, na.rm = TRUE),
    #Count = ~sum(!is.na(.x))
  ), .names = "{.col}_{.fn}"))
summary_stats
summary_stats %>% melt(.) %>% mutate(value = round(value, 2))


anova_results


```
### Plot distribution of z-transformed non-normal cognitive scores
```{r}
cog_test_scores%>% dplyr::select(where(is.numeric)) %>% test_normality() > 0.05 # none of the scores are normal (when BD and HC are included) 
cog_test_scores
```

This section identifies and removes extreme outliers (more than 3 SD from the mean) from cognitive test scores using a standard deviation-based filter. It then:    
1. Visualizes the distributions of the cleaned variables.  
2. Computes summary statistics (mean and SD) for each variable.   
3. Tests for normality on numeric variables to determine suitability for parametric analyses.
```{r}
# there 5 subjects with scores > 3 std devs above the mean

# Remove extreme outliers (beyond ±3 standard deviations) and round remaining values
outliers_removed <- cog_test_scores %>%
  dplyr::select(-trvscerw, -trvsoerw) %>%  # Drop specific variables that are ranked or binary
  mutate_if(is.numeric, ~ ifelse(abs(. - mean(.)) / sd(.) <= 2, ., NA)) %>%
  dplyr::select_if(~!all(is.na(.))) %>%    # Drop columns that are all NA
  mutate_if(is.numeric, ~ round(., 3))

# Visualize cleaned distributions with histograms faceted by variable
outliers_removed %>% filter(stdyptid != 'BD051') %>%
  reshape2::melt() %>%
  ggplot(aes(x = value, fill = dxgroupFK)) +
  geom_histogram() +
  facet_grid(~ variable, scales = "free") +
  theme_bw()


# Optional: Check which numeric variables are approximately normally distributed (p > 0.05)
# Requires a custom function `test_normality()` applied across columns
outliers_removed %>%
  select_if(is.numeric) %>%
  test_normality > 0.05 


anova_results <- outliers_removed %>%
  pivot_longer(
    cols      = -c(stdyptid, gender, dxgroupFK),
    names_to  = "test",
    values_to = "score"
  ) %>%
  filter(!is.na(score)) %>%
  group_by(test) %>%
  do(broom::tidy(aov(score ~ factor(dxgroupFK), data = .))) %>%
  filter(term == "factor(dxgroupFK)") %>%
  # explicitly call dplyr::select, then rename
  dplyr::select(test, statistic, p.value, df) %>%
  rename(
    Test    = test,
    F_value = statistic,
    p_value = p.value,
    Df      = df
  ) %>%
  arrange(p_value)
anova_results
ggsave("output/figures/cog_scores_outlier_removed_distribution.jpeg",dpi = 500, width = 15, height = 5)
dev.off()
```


This script performs a PCA-based dimensionality reduction on cognitive task variables after removing outliers. It includes the following steps:

1. Adjusts skewed values by adding a constant to ensure all values are positive.
2. Removes variables that contain only NAs.
3. Applies a log10 transformation and standardizes numeric variables.
4. Flip signs for specific variables so that higher values reflect better cognitive performance.
5. Conducts Principal Component Analysis (PCA) and extracts the first principal component.
6. Merges the PCA scores with the original dataframe containing additional variables like age or diagnosis
```{r}
# Add a constant to shift all values of 'tmtb_a' to be positive
constant <- abs(min(outliers_removed$tmtb_a, na.rm = TRUE)) + 1
outliers_removed$tmtb_a <- outliers_removed$tmtb_a + constant

# Remove columns that are entirely NA
outliers_removed <- outliers_removed %>%
  select_if(function(x) !all(is.na(x)))
outliers_removed

# Perform PCA on numeric variables after log transformation and scaling
pca_result <- outliers_removed %>%
  select_if(is.numeric) %>%
  log10() %>%
  # we flip the sign for these scores so that higher=worse performance
  mutate_if(is.numeric, scale) %>%
  mutate(
    cwicnrw = -cwicnrw,
    cwiwrrw = -cwiwrrw,
    cwiirw = -cwiirw,
    cwiisrw = -cwiisrw
  ) %>%
  principal(nfactors =2 )

pca_result$loadings
# Create a dataframe that includes non-cognitive variables (e.g., age, diagnosis)
outliers_removed_df <- df %>% 
  filter(stdyptid %in% outliers_removed$stdyptid)
# Enter loadings manually

 
# Bind PCA scores to the full dataframe and flip direction so that higher = better performance
outliers_removed_df_pc <- cbind(as.data.frame(outliers_removed_df), pca_result$scores*-1)
# View the final dataframe with PCA scores
outliers_removed_df_pc


 
# Compute mean and standard deviation for each numeric variable
summary_stats <- outliers_removed_df_pc%>% filter(stdyptid != 'BD051') %>%
  dplyr::select( -stdyptid, -gender) %>% group_by(dxgroupFK) %>%
  summarise(across(everything(), list(
    Mean = ~mean(.x, na.rm = TRUE),
    SD = ~sd(.x, na.rm = TRUE)
  ), .names = "{.col}_{.fn}"))

# Reshape and round summary statistics for easier inspection
summary_stats %>%
  reshape2::melt() %>%
  mutate(value = round(value, 2))


# Flip loadings and scores if you want higher PC1 = better performance
loadings_df <- as.data.frame(unclass(pca_result$loadings)) %>%
  tibble::rownames_to_column("variable") %>%
  rename(loading = RC2) %>%
  mutate(loading = -loading)

ggplot(loadings_df, aes(x = reorder(variable, loading), y = loading)) +
  geom_col(width = 0.7) +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_text(aes(label = round(loading, 2)),
            hjust = ifelse(loadings_df$loading > 0, -0.15, 1.15),
            size = 4) +
  expand_limits(y = c(min(loadings_df$loading) - 0.15,
                      max(loadings_df$loading) + 0.15)) +
  labs(
    title = "PC2 loadings",
    x = NULL,
    y = "Loading"
  ) +
  theme_minimal(base_size = 13)
 

 
 
```


### cleaned dataset used for examining relationship between cognitive performance (PC1) and residualized brain-PADs from all four models
```{r}
#write.csv(outliers_removed_df_pc, 'data_processed/outliers_removed_cognitive_test_scores.csv')
```

### Scree Plot
```{r}
# Extract eigenvalues from the PCA result
eigenvalues <- pca_result$values
pca_result
# Calculate the proportion of variance explained by each principal component
variance_explained <- eigenvalues / sum(eigenvalues)

# Create a scree plot
plot(1:length(variance_explained), variance_explained, type = "b", 
     xlab = "Principal Component", ylab = "Proportion of Variance Explained",
     main = "Scree Plot")
variance_explained
```



# ANALYSIS
# Linear regression model with bpad as outcome variable
### Ridge Bpad
```{r}
# Filter out rows with missing bpad_ridge
filtered_df <- outliers_removed_df_pc %>%
  filter(complete.cases(bpad_ridge))




resid_mod <- lm(bpad_ridge ~   gender*agevisit, data = filtered_df)
resid_bpad_ridge <- resid(resid_mod)
summary(resid_mod)

# Regress PC1 on residualized bpad_photon
pca_cog_mod <- lm(RC2 ~ resid_bpad_ridge*dxgroupFK, data = filtered_df)
summary(pca_cog_mod)

# age of HC vs BD for multimodal
age_filtered_aov <- aov(agevisit ~ factor(dxgroupFK), data = filtered_df)
summary(age_filtered_aov)

summary(pca_cog_mod)

# Get partial eta squared
getPartialEtaSq(pca_cog_mod)


pca_cog_hp_mod <- lm(PC1 ~  totalGMvolume , data = filtered_df)
summary(pca_cog_hp_mod)
getPartialEtaSq(pca_cog_hp_mod) 

pca_cog_hp_mod <- lm(PC1 ~  totalGMvolume , data = out)
summary(pca_cog_hp_mod)
getPartialEtaSq(pca_cog_hp_mod) 




```

### PHOTON Bpad
```{r}
resid_mod <- lm(bpad_photon ~   gender*agevisit, data = outliers_removed_df_pc)
resid_bpad_photon <- resid(resid_mod)
summary(resid_mod)
pca_cog_mod <- lm(PC1 ~  resid_bpad_photon*dxgroupFK  , data = outliers_removed_df_pc)
summary(pca_cog_mod)
getPartialEtaSq(pca_cog_mod) 


pca_cog_hp_mod <- lm(PC1 ~  resid_bpad_photon*HistoryPsychosis , data = outliers_removed_df_pc)
summary(pca_cog_hp_mod)
getPartialEtaSq(pca_cog_hp_mod) 
 



```


### BrainAgeR Bpad
```{r}
resid_mod <- lm(bpad_BrainAgeR ~   gender*agevisit, data = outliers_removed_df_pc)
resid_bpad_BrainAgeR <- resid(resid_mod)
grubbs.test(resid_bpad_BrainAgeR)
summary(resid_mod)
pca_cog_mod <- lm(RC1 ~  resid_bpad_BrainAgeR*dxgroupFK , data = outliers_removed_df_pc)
summary(pca_cog_mod)
getPartialEtaSq(pca_cog_mod) 


pca_cog_hp_mod <- lm(RC1 ~  resid_bpad_BrainAgeR*HistoryPsychosis , data = outliers_removed_df_pc)
summary(pca_cog_hp_mod)
getPartialEtaSq(pca_cog_hp_mod) 



 
```

```{r}
resid_mod <- lm(bpad_cnn ~   gender*agevisit, data = outliers_removed_df_pc)
resid_bpad_cnn <- resid(resid_mod)
grubbs.test(resid_bpad_cnn)
 
pca_cog_mod <- lm(RC1 ~ resid_bpad_cnn*dxgroupFK, data = outliers_removed_df_pc)
summary(pca_cog_mod) 
getPartialEtaSq(pca_cog_mod) 
 




pca_cog_hp_mod <- lm(PC1 ~  resid_bpad_cnn*HistoryPsychosis , data = outliers_removed_df_pc)
summary(pca_cog_hp_mod)
getPartialEtaSq(pca_cog_hp_mod) 
``` 

### Four subplots showing the linear relationship between resid bpad and PC1

```{r fig.width=5, fig.height=5}
# Calculate R-squared for the entire dataset
r2_values_BrainAgeR <- summary(lm(RC1 ~ resid_bpad_BrainAgeR, data = outliers_removed_df_pc))$r.squared


# Plot with CI only for overall regression line
base_plot <- ggplot(outliers_removed_df_pc, aes(x = PC1, y = resid_bpad_BrainAgeR, color = factor(dxgroupFK))) + geom_point(alpha = 0.6) +
  
  # Group-specific regression lines WITHOUT confidence intervals
  geom_smooth(
    aes(group = factor(dxgroupFK), color = factor(dxgroupFK)),
    method = "lm", formula = y ~ x, se = FALSE
  ) + # Overall regression line WITH confidence interval
  geom_smooth(
    method = "lm", formula = y ~ x, se = TRUE,
    color = 'darkorchid2', fill = 'darkorchid2', alpha = 0.1
  ) +  scale_color_manual(
    name = "Group",
    labels = c("HC", "BD"),
    values = c('darkblue', 'red')
  ) +
  theme_bw() + theme(
    text = element_text(size = 11),
    legend.position = "bottom"
  ) +
  coord_cartesian(xlim = c(-20, 20), ylim = c(-3, 3)) +
  labs(
    x = "Residualized Brain-PAD (BrainageR)", y = "Executive Functioning Component",
  ) 

# Show the plot
print(base_plot)


# 
jpeg('output/figures/pc1_brainager.jpeg', width = 4000, height = 4000, units = "px", res = 500)
base_plot
dev.off()


# Annotate the plot with R-squared values for the first variable
plot1 <- base_plot + geom_richtext(aes(label = paste0("R<sup>2</sup> = ", format(round(r2_values_BrainAgeR, digits=3), nsmall = 2) ), x = Inf, y = -Inf, hjust = 1, vjust = -0.5), color = 'black') 
jpeg('output/figures/lm_PC1_resid_bpad_BrainAgeR.jpeg', width = 4000, height = 4000, units = "px", res = 500)
plot1
dev.off()
```


```{r fig.width=5, fig.height=5}
base_plot <- ggplot(outliers_removed_df_pc, aes(x = resid_bpad_photon, y = PC1,  color = factor(dxgroupFK)))   + geom_point(alpha = 0.6) +
  
  # Group-specific regression lines WITHOUT confidence intervals
  geom_smooth(
    aes(group = factor(dxgroupFK), color = factor(dxgroupFK)),
    method = "lm", formula = y ~ x, se = FALSE
  ) + # Overall regression line WITH confidence interval
  geom_smooth(
    method = "lm", formula = y ~ x, se = TRUE,
    color = 'darkorchid2', fill = 'darkorchid2', alpha = 0.1
  ) +  scale_color_manual(
    name = "Group",
    labels = c("HC", "BD"),
    values = c('darkblue', 'red')
  ) +
  theme_bw() + theme(
    text = element_text(size = 11),
    legend.position = "bottom"
  ) +
  coord_cartesian(xlim = c(-20, 20), ylim = c(-3, 3)) +
  labs(
    x = "Residualized Brain-PAD (PHOTON)", y = "Executive Functioning Component",
  ) 


# 
jpeg('output/figures/pc1_photon.jpeg', width = 4000, height = 4000, units = "px", res = 500)
base_plot
dev.off()

plot2 <- base_plot +
# geom_richtext(aes(label = paste0("R<sup>2</sup> = ",  format(round(r2_values_photon, digits=3), nsmall = 2) ), x = Inf, y = -Inf, hjust = 1, vjust = -0.5), color = 'black') +
  labs(x = "Residual Brain-PAD",title = "PHOTON-BA")
plot2
#outliers_removed_df_pc

 

# Calculate R and p-value
model_photon <- lm(PC1 ~ resid_bpad_photon, data = outliers_removed_df_pc)
r_value <- summary(model_photon)$r.squared^(1/2)  # R value
p_value <- summary(model_photon)$coefficients[2, 4]  # p-value for resid_bpad_photon

base_plot #+geom_richtext(aes(label = paste0("R = ", format(round(r_value, digits=3), nsmall = 2), 
                                  #", p = ", format.pval(p_value) ), x = Inf, y = -Inf, hjust = 1, vjust = -0.5), color = 'black') +
#  labs(x = "Residualized Brain-PAD")

#jpeg('output/figures/lm_PC1_resid_bpad_photon.jpeg', width = 4000, height = 4000, units = "px", res = 500)
plot2
#dev.off()
```
 

```{r}

base_plot <- ggplot(outliers_removed_df_pc, aes(x = resid_bpad_cnn, y = PC1, color = factor(dxgroupFK)))   + geom_point(alpha = 0.6) +
  
  # Group-specific regression lines WITHOUT confidence intervals
  geom_smooth(
    aes(group = factor(dxgroupFK), color = factor(dxgroupFK)),
    method = "lm", formula = y ~ x, se = FALSE
  ) + # Overall regression line WITH confidence interval
  geom_smooth(
    method = "lm", formula = y ~ x, se = TRUE,
    color = 'darkorchid2', fill = 'darkorchid2', alpha = 0.1
  ) +  scale_color_manual(
    name = "Group",
    labels = c("HC", "BD"),
    values = c('darkblue', 'red')
  ) +theme_bw() + theme(
    text = element_text(size = 11),
    legend.position = "bottom"
  ) +coord_cartesian(xlim = c(-20, 20), ylim = c(-3, 3)) +
  labs(
    x = "Residualized Brain-PAD (DenseNet)", y = "Executive Functioning Component",
  ) 


jpeg('output/figures/pc1_densenet.jpeg', width = 4000, height = 4000, units = "px", res = 500)
base_plot
dev.off()



r2_values_cnn <- summary(lm(PC1 ~ resid_bpad_cnn, data = outliers_removed_df_pc))$r.squared

jpeg('output/figures/lm_PC1_resid_bpad_cnn.jpeg', width = 4000, height = 4000, units = "px", res = 500)
plot3 <- base_plot +  geom_richtext(aes(label = paste0("R<sup>2</sup> = ", format(round(r2_values_cnn, digits=3), nsmall = 2) ), x = Inf, y = -Inf, hjust = 1, vjust = -0.5), color = 'black') + labs(x = "Residual Brain-PAD",title = "MIDI")
plot3
dev.off() 

```

```{r}

base_plot <- ggplot(filtered_df, aes(x = resid_bpad_ridge, y = PC1, color = factor(dxgroupFK)))  + geom_point(alpha = 0.6) +
  
  # Group-specific regression lines WITHOUT confidence intervals
  geom_smooth(
    aes(group = factor(dxgroupFK), color = factor(dxgroupFK)),
    method = "lm", formula = y ~ x, se = FALSE
  ) + # Overall regression line WITH confidence interval
  geom_smooth(
    method = "lm", formula = y ~ x, se = TRUE,
    color = 'darkorchid2', fill = 'darkorchid2', alpha = 0.1
  ) +  scale_color_manual(
    name = "Group",
    labels = c("HC", "BD"),
    values = c('darkblue', 'red')
  ) +theme_bw() + theme(
    text = element_text(size = 11),
    legend.position = "bottom"
  ) +coord_cartesian(xlim = c(-20, 20), ylim = c(-3, 3)) +
  labs(
    x = "Residualized Brain-PAD (Multimodal)", y = "Executive Functioning Component",
  ) 


jpeg('output/figures/pc1_multimodal_ridge.jpeg', width = 4000, height = 4000, units = "px", res = 500)
base_plot
dev.off()


r2_values_ridge <- summary(lm(PC1 ~ resid_bpad_ridge, data = filtered_df ))$r.squared
plot4 <- base_plot +
  geom_richtext(aes(label = paste0("R<sup>2</sup> = ", format(round(r2_values_ridge, digits=3), nsmall = 2) ), x = Inf, y = -Inf, hjust = 1, vjust = -0.5), color = 'black') +
  labs(x = "Residual Brain-PAD", title = 'UCSD Multimodal')

jpeg('output/figures/lm_PC1_resid_bpad_ridge.jpeg', width = 4000, height = 4000, units = "px", res = 500)
 
plot4
dev.off() 
```


### Log transform cog scores
```{r}
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3577045/ PCA to run on DKEFS
# Ensure constant is computed properly
constant <- abs(min(outliers_removed$tmtb_a, na.rm = TRUE)) + 1

# Shift tmtb_a to be all positive
outliers_removed$tmtb_a <- outliers_removed$tmtb_a + constant

log_transformed


# Add a constant to the scores to make them positive
log_transformed <- outliers_removed %>%  mutate_if(., is.numeric, log10)

log_transformed %>% 
  dplyr::select(where(is.numeric)) %>%
  dplyr::select(where(~ sum(!is.na(.)) >= 3)) %>%
  summarise(across(everything(), ~shapiro.test(.)$p.value))


log_transformed %>% filter(stdyptid != 'BD051') %>%reshape2::melt(.) %>%ggplot(., aes(x = value, fill=dxgroupFK)) +
  geom_histogram() +
  facet_wrap(~ variable, scales = "free")  +
    scale_fill_discrete(
      name   = "Group",
      labels = c("HC", "BD")) + theme_classic() + labs(x = "Log₁₀(Value)", y = "Count")

ggsave("output/figures/log_transformed_pc_1_cog_scores.jpeg",dpi = 500, width = 15, height = 10)
dev.off()

log_transformed #%>% group_by(dxgroupFK) %>% tally()


log_transformed
# Compute mean and standard deviation for each numeric variable
summary_stats <- log_transformed %>%
  dplyr::select( -stdyptid, -gender) %>% group_by(dxgroupFK) %>%
  summarise(across(everything(), list(
    Mean = ~mean(.x, na.rm = TRUE),
    SD = ~sd(.x, na.rm = TRUE)
  ), .names = "{.col}_{.fn}"))
summary_stats
# Reshape and round summary statistics for easier inspection
summary_stats %>%
  reshape2::melt() %>%
  mutate(value = round(value, 2))

log_transformed %>% group_by(dxgroupFK) %>% tally()
log_transformed
log_transformed %>%
  pivot_longer(
    cols      = -c(stdyptid, gender, dxgroupFK),
    names_to  = "test",
    values_to = "score"
  ) %>%
  filter(!is.na(score)) %>%
  group_by(test) %>%
  do(broom::tidy(aov(score ~ factor(dxgroupFK), data = .))) %>%
  filter(term == "factor(dxgroupFK)") %>%
  
  # explicitly call dplyr::select, then rename
dplyr::select(test, statistic, p.value, df) %>%
  rename(
    Test    = test,
    F_value = statistic,
    p_value = p.value,
    Df      = df
  ) %>% arrange(p_value)


# what is the comparison for PC1
pc1_compare <- aov(PC1 ~ factor(dxgroupFK), data = outliers_removed_df_pc)
summary(pc1_compare)

outliers_removed_df_pc %>% group_by(dxgroupFK) %>%
  summarise(across(PC1, list(
    Mean = ~mean(.x, na.rm = TRUE),
    SD = ~sd(.x, na.rm = TRUE)
  ), .names = "{.col}_{.fn}"))
```


### Summary stats for HC = 22
```{r}
filtered_df_log <- log_transformed[
  log_transformed$stdyptid %in% filtered_df$stdyptid]

filtered_df_log %>%
  dplyr::select( -stdyptid, -gender) %>% group_by(dxgroupFK) %>%
  summarise(across(everything(), list(
    Mean = ~mean(.x, na.rm = TRUE),
    SD = ~sd(.x, na.rm = TRUE)
  ), .names = "{.col}_{.fn}"))  %>%
  reshape2::melt() %>%
  mutate(value = round(value, 2))

summary_filtered_df
summary_filtered_df <- filtered_df_log %>%
  pivot_longer(
    cols      = -c(stdyptid, gender, dxgroupFK),
    names_to  = "test",
    values_to = "score"
  ) %>%
  filter(!is.na(score)) %>%
  group_by(test) %>%
  do(broom::tidy(aov(score ~ factor(dxgroupFK), data = .))) %>%
  filter(term == "factor(dxgroupFK)") %>%
  
  # explicitly call dplyr::select, then rename
dplyr::select(test, statistic, p.value, df) %>%
  rename(
    Test    = test,
    F_value = statistic,
    p_value = p.value,
    Df      = df
  ) %>% arrange(p_value)




filtered_df_pc1 <- outliers_removed_df_pc[
  outliers_removed_df_pc$stdyptid %in% filtered_df$stdyptid,]

# With only
filtered_df_pc1 %>% group_by(dxgroupFK)  %>% dplyr::select(PC1) %>%
  summarise(across(everything(), list(
    Mean = ~mean(.x, na.rm = TRUE),
    SD = ~sd(.x, na.rm = TRUE)
  ), .names = "{.col}_{.fn}"))

pc1_compare <- aov(PC1 ~ factor(dxgroupFK), data = filtered_df_pc1)
summary(pc1_compare)
```


