---
title: "compare_four_models"
author: "Hui Xin Ng"
date: "2023-03-19"
output: html_document
---

# Load required libraries
```{r setup, include=FALSE}
library(dplyr)
library(tidyr)       # used with dplyr pipelines
library(ggplot2)
library(reshape2)    # for melt() – even though deprecated, you're still using it
library(readxl)      # if reading Excel files, although not shown in session
library(haven)       # for `read_sav()` when loading SPSS files
library(DescTools)   # for `icc()` function
library(data.table)  # used with `setDT()` and `melt()`
library(lme4)        # for `lmer()`
library(lmerTest)    # for p-values in `lmer`
library(lmtest)      # optional, used for model testing
library(irr)         # for inter-rater reliability (e.g. `icc()`, alternative to DescTools)

```

# What does this rmd file accomplish?
Visualization: 
- Density plots to assess the distribution of Brain-PAD values and predicted brain ages across models
- Scatter plots to explore the relationship between chronological age and brain age predictions.

Normality Testing: 
Shapiro-Wilk tests to evaluate the normality of Brain-PAD and predicted age distributions across different models and groups.

Correlation Analysis:
Correlations between chronological age and Brain-PAD values, as well as between predicted ages across models, including separate analyses for the HC and BD groups. The results are presented in correlation matrices and heatmaps.

Standard dev: SD of Brain-PAD and predicted age values as measures of variance within both HC and BD groups.
 
# Load and clean original dataset (Master_BAI.....) but we may not use it
```{r original df}
# this is ocassionally used for sanity checks
path = file.path("../data_original/Master_BAI_080720_with_051517completers_060117imagingmaster_genotypes_brainages_ENIGMA_FS_MERGEFK_30122020.sav")
original_df = read_sav(path)
original_df <- original_df %>% dplyr::select("trnsctrw", "cwiisnrs","cwiisis", "cwiicns","fsrp5","stdyptid", 'visit', 'dxgroupFK', 'agevisit') %>% filter(!is.na(trnsctrw)|!is.na(fsrp5)) %>% arrange(stdyptid)
original_df <- setDT(original_df)[, lapply(.SD, na.omit), by = stdyptid]
```

# full dataset with all brain ages
```{r full dataset from collate_all_brainages.ipynb}
df<-read.csv("../data_processed/df_all_brainage_2025.csv") 
df %>% group_by(dxgroupFK) %>%tally()

```

# Create derived variables including Brain-PAD (predicted age - chronological age)
```{r}
df  <- df %>% mutate(Age2 = agevisit**2, Age3 = agevisit**3, 
                     bpad_ridge = ridge_BrainAge - agevisit,
                     bpad_photon = age_prediction - agevisit,
                     bpad_BrainAgeR = ageprediction_BrainAgeR - agevisit,
                     bpad_cnn = cnn_BrainAge - agevisit)


```

# Helper functions
```{r}
# Function to filter out data points outside of 3 SDs
filter_outliers <- function(x) {
  sd_x <- sd(x)
  mean_x <- mean(x)
  upper_limit <- mean_x + 3 * sd_x
  lower_limit <- mean_x - 3 * sd_x
  return(between(x, lower_limit, upper_limit))
}

# Function to plot distribution and run Shapiro-Wilk test, Apply function to each Brain-PAD variable by group
plot_and_test <- function(df, variable, plot_type = "density", fill_color = NULL, legend_labels = NULL, legend_title = "Group", x_label = NULL, y_label = NULL) 
{ 
 # Create plot
  p <- ggplot(df, aes(x = .data[[variable]], fill = as.factor(df[[fill_color]]))) + 
    geom_density(alpha = 0.3) + coord_cartesian(ylim=c(0, 0.075))+
    labs(fill = legend_title, x = x_label, y = y_label) + 
    scale_fill_manual(values = c("darkgrey", "red"), labels = legend_labels) +
    theme_classic()
print(p) # Print plot
  normality_test <- shapiro.test(df[[variable]])   # Perform Shapiro-Wilk test for normality
  print(paste("Shapiro-Wilk p-value for", variable, ":", normality_test$p.value))
}

```

# Distribution and Normality Test of Brain-PAD 
```{r}

plot_and_test(df, "agevisit", fill_color = "dxgroupFK", legend_labels = c("HC", "BD"), legend_title = "Group", x= 'Chronological Age', y='Density')


df <- df %>%
  filter(filter_outliers(bpad_ridge),
         filter_outliers(bpad_BrainAgeR),
         filter_outliers(bpad_photon),
         filter_outliers(bpad_cnn))

# Example usage of the function for your variables
plot_and_test(df, "bpad_ridge", fill_color = "dxgroupFK", legend_labels = c("HC", "BD"), legend_title = "Group", x= 'Multimodal Brain-PAD', y='Density')
ggsave("output/figures/density_plot_bpad_ridge.jpeg",dpi = 500, width = 5, height = 5)
dev.off()

# Example usage of the function for your variables
plot_and_test(df, "bpad_photon", fill_color = "dxgroupFK", legend_labels = c("HC", "BD"), legend_title = "Group", x= 'PHOTON Brain-PAD', y='Density')
ggsave("output/figures/density_plot_bpad_photon.jpeg",dpi = 500, width = 5, height = 5)
dev.off()

# Example usage of the function for your variables
plot_and_test(df, "bpad_BrainAgeR", fill_color = "dxgroupFK", legend_labels = c("HC", "BD"), legend_title = "Group", x= 'BrainageR Brain-PAD', y='Density')
ggsave("output/figures/density_plot_bpad_BrainAgeR.jpeg",dpi = 500, width = 5, height = 5)
dev.off()
 
 # Example usage of the function for your variables
plot_and_test(df, "bpad_cnn", fill_color = "dxgroupFK", legend_labels = c("HC", "BD"), legend_title = "Group", x= 'DenseNet Brain-PAD', y='Density')
ggsave("output/figures/density_plot_bpad_cnn.jpeg",dpi = 500,  width = 5, height = 5)
dev.off()

 # List of Brain-PAD columns
bpad_cols <- c("bpad_cnn", "bpad_photon", "bpad_ridge", "bpad_BrainAgeR")

# Generate all pairwise combinations
combos <- combn(bpad_cols, 2, simplify = FALSE)

# Run KS test for each pair and store results
ks_results <- lapply(combos, function(pair) {
  test <- ks.test(df[[pair[1]]], df[[pair[2]]])
  list(
    model_1 = pair[1],
    model_2 = pair[2],
    statistic = test$statistic,
    p_value = test$p.value
  )
})

# Convert to data frame
ks_df <- do.call(rbind, lapply(ks_results, as.data.frame))
# View results
ks_df
 
```
# Distribution and Normality Test of   Predicted Brain Ages 
```{r}

# ---- Distribution and Normality Test of Brain Age -----  
# Example usage of the function for your variables
plot_and_test(df, "ridge_BrainAge", fill_color = "dxgroupFK", legend_labels = c("HC", "BD"), legend_title = "Group", x= 'Predicted Age  (Multimodal)', y='Density')
#ggsave("output/figures/density_plot_ridge_BrainAge.jpeg",dpi = 500, width = 12, height = 3)
dev.off()
# Example usage of the function for your variables
plot_and_test(df, "age_prediction", fill_color = "dxgroupFK", legend_labels = c("HC", "BD"), legend_title = "Group", x= 'Predicted Age  (PHOTON-AI)', y='Density')
#ggsave("output/figures/density_plot_photon_BrainAge.jpeg",dpi = 500, width = 12, height = 3)
dev.off()

# Example usage of the function for your variables
plot_and_test(df, "ageprediction_BrainAgeR", fill_color = "dxgroupFK", legend_labels = c("HC", "BD"), legend_title = "Group", x= 'Predicted Age  (BrainageR)', y='Density')
#ggsave("output/figures/density_plot_predicted_BrainAgeR.jpeg",dpi = 500, width = 12, height = 3)
dev.off() 
 # Example usage of the function for your variables
plot_and_test(df, "cnn_BrainAge", fill_color = "dxgroupFK", legend_labels = c("HC", "BD"), legend_title = "Group", x= 'Predicted Age (MIDI)', y='Density')
ggsave("output/figures/density_plot_cnn_BrainAge.jpeg",dpi = 500, width = 5, height = 5)
dev.off()
```

# correlation between chronological age and brain-pad
```{r}
corr_age_brainpad_across_models <- df %>% dplyr::select(bpad_ridge, bpad_photon , bpad_BrainAgeR ,bpad_cnn, dxgroupFK, gender, agevisit) %>%  melt(id = c('agevisit', 'dxgroupFK', 'gender'))%>% group_by(variable, dxgroupFK) %>% summarise(pval = as.numeric(cor.test(agevisit, value)[3]),                                                                correlation = cor(agevisit, value))%>% mutate(pval = round(pval, 3), correlation = round(correlation, 3))

df

corr_age_brainpad_across_models
# corr_age_brainpad_across_models %>%
#   write.csv("output/data/corr_age_brainpad_across_models.csv", row.names = FALSE)
```

# correlation between chronological age and predicted age
```{r}
corr_age_predictedage_across_models <- df %>% dplyr::select(ridge_BrainAge, ageprediction_BrainAgeR , age_prediction ,cnn_BrainAge, dxgroupFK, gender, agevisit) %>%  melt(id = c('agevisit', 'dxgroupFK', 'gender'))%>% group_by(variable, dxgroupFK) %>% summarise(pval = as.numeric(cor.test(agevisit, value)[3]),  correlation = cor(agevisit, value))%>% mutate(pval = round(pval, 3), correlation = round(correlation, 3)) %>% filter()
corr_age_predictedage_across_models
#corr_age_predictedage_across_models %>% write.csv(., 'output/data/corr_age_predictedage_across_models.csv')
```
# Correlation between brain-PAD and heatmap of matrix
```{r bpad_corr_heatmap.jpeg}
# Select the columns you want to calculate pairwise correlations 
# Calculate pairwise correlations
library(reshape2)

correlation_matrix <- df %>%
  dplyr::select(bpad_photon, bpad_BrainAgeR, bpad_cnn, bpad_ridge) %>%  # reorder here
  cor(.)

write.csv(round(correlation_matrix, 3),
          "output/data/bpad_corr_matrix.csv",
          row.names = TRUE)

# Melt the matrix
corr_df <- melt(correlation_matrix)

# Ensure factor ordering matches PHOTON → BrainageR → DenseNet → Multimodal
corr_df$Var1 <- factor(corr_df$Var1,
                       levels = c("bpad_photon", "bpad_BrainAgeR", "bpad_cnn", "bpad_ridge"))
corr_df$Var2 <- factor(corr_df$Var2,
                       levels = c("bpad_photon", "bpad_BrainAgeR", "bpad_cnn", "bpad_ridge"))

# Plot heatmap
bpad_corr_heatmap <- ggplot(corr_df, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "white", high = "brown", midpoint = 0) +
  geom_text(aes(label = sprintf("%.2f", value)), color = "black", size = 4) +
  theme_minimal() +
  labs(title = "Brain-PAD Correlation Heatmap across algorithms", x = "", y = "") +
  scale_x_discrete(
    labels = c(
      "bpad_photon" = "PHOTON",
      "bpad_BrainAgeR" = "BrainageR",
      "bpad_cnn" = "DenseNet",
      "bpad_ridge" = "Multimodal")) +
  scale_y_discrete(
    labels = c(
      "bpad_photon" = "PHOTON",
      "bpad_BrainAgeR" = "BrainageR",
      "bpad_cnn" = "DenseNet",
      "bpad_ridge" = "Multimodal")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

bpad_corr_heatmap

ggsave("output/figures/bpad_corr_heatmap.jpeg",
       width = 8, height = 6, dpi = 300)
dev.off()


```

# Correlation between brain-PAD in HC across models and heatmap of corr matrix
```{r}
corr_hc_bpad <- df %>% dplyr::filter(dxgroupFK == 0) %>% dplyr::select(bpad_ridge, bpad_photon , bpad_BrainAgeR ,bpad_cnn)  %>% cor(.)    


# Melt the matrix for ggplot
hc_bpad_corr_heatmap <- corr_hc_bpad  %>% melt(.) %>%
  mutate(
    Var1 = factor(Var1, levels = rownames(corr_hc_bpad), labels = custom_labels),
    Var2 = factor(Var2, levels = rownames(corr_hc_bpad), labels = custom_labels)
  ) %>% ggplot(., aes(x = Var1, y =Var2, fill=value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "white",   high = "lightblue", midpoint = 0) +
  geom_text(aes(label = sprintf("%.2f", value)), color = "black", size = 4) +  
  theme_minimal() +
  labs(title = "Brain-PAD correlation Heatmap across algorithms in HC", x = "", y = "") +
  # Rename the x- and y-axis labels here
  scale_x_discrete(
    labels = c(
      "bpad_ridge" = "Multimodal Ridge",
      "bpad_photon" = "PHOTON",
      "bpad_BrainAgeR" = "BrainageR",
      "bpad_cnn" = "DenseNet MIDI"
    )
  ) +
  scale_y_discrete(
    labels = c(
      "bpad_ridge" = "Multimodal Ridge",
      "bpad_photon" = "PHOTON",
      "bpad_BrainAgeR" = "BrainageR",
      "bpad_cnn" = "DenseNet MIDI"
    )
  ) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
  )

hc_bpad_corr_heatmap
ggsave("output/figures/hc_bpad_corr_heatmap.jpeg",
  width = 8, height = 6, dpi = 300)
dev.off() 


```

# Standard deviation as a measure of variance in Brain-PAD in BD and HC groups
```{r}
df %>% dplyr::select(bpad_ridge, bpad_photon , bpad_BrainAgeR ,bpad_cnn, dxgroupFK, gender, agevisit, ridge_BrainAge, age_prediction , cnn_BrainAge ,ageprediction_BrainAgeR) %>% group_by(dxgroupFK) %>% 
  summarise(sd_bpad_photon = round(sd(bpad_photon), 2), 
            sd_bpad_BrainAgeR = round(sd(bpad_BrainAgeR), 2), 
            sd_bpad_cnn = round(sd(bpad_cnn), 2), 
            sd_bpad_ridge = round(sd(bpad_ridge), 2), 
                # Raw predicted brain age metrics
     sd_ridge_BrainAge = round(sd(ridge_BrainAge, na.rm = TRUE), 2),
     sd_age_prediction = round(sd(age_prediction, na.rm = TRUE), 2),
     sd_cnn_BrainAge = round(sd(cnn_BrainAge, na.rm = TRUE), 2),
     sd_ageprediction_BrainAgeR = round(sd(ageprediction_BrainAgeR, na.rm = TRUE), 2),
     sd_agevisit = round(sd(agevisit, na.rm = TRUE), 2)) # Visit age

```

# Correlation between predicted ages
```{r}
# --- Correlation Matrix Brain Age ---- 
# Select the columns you want to calculate pairwise correlations for
cor_data <-  df %>%  filter(dxgroupFK == 0)%>% dplyr::select(ridge_BrainAge, age_prediction , cnn_BrainAge ,ageprediction_BrainAgeR)  %>% cor(.)
 
# Calculate pairwise correlations
correlation_matrix <- cor(cor_data)

# Print correlation matrix
round(correlation_matrix, 3)

write.csv(round(correlation_matrix, 3), "output/data/predicted_ages_corr_matrix.csv", row.names = TRUE)

df# Select the columns you want to calculate pairwise correlations for
cor_data <- df %>% dplyr::select( age_prediction , cnn_BrainAge ,ageprediction_BrainAgeR, agevisit)  %>% cor(.)
round(cor_data,3)


df# Select the columns you want to calculate pairwise correlations for
cor_data <- df %>% dplyr::select( ridge_BrainAge, age_prediction , cnn_BrainAge ,ageprediction_BrainAgeR, agevisit) %>% filter(!is.na(ridge_BrainAge)) %>% cor(.)
round(cor_data^2,3)

# Calculate pairwise correlations
correlation_matrix <- cor(cor_data)

# Print correlation matrix
round(correlation_matrix, 3)
```

# Correlation between predicted ages and heatmap of matrix
```{r bpad_corr_heatmap.jpeg}
# Select the columns you want to calculate pairwise correlations 
# Calculate pairwise correlations
correlation_matrix <-  df %>% dplyr::select(ridge_BrainAge, age_prediction , cnn_BrainAge ,ageprediction_BrainAgeR)  %>% cor(.) 
correlation_matrix
custom_labels <- c("Multimodel Ridge", "PHOTON",  "DenseNet MIDI", 'BrainageR')

# Melt the matrix for ggplot
predicted_ages_corr_heatmap <- correlation_matrix %>% melt(.) %>%
  mutate(
    Var1 = factor(Var1, levels = rownames(correlation_matrix), labels = custom_labels),
    Var2 = factor(Var2, levels = rownames(correlation_matrix), labels = custom_labels)
  ) %>%ggplot(., aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "white",   high = "brown", midpoint = 0) +
  geom_text(aes(label = sprintf("%.2f", value)), color = "black", size = 4) +
  theme_minimal() +
  labs(title = "Predicted Age correlation heatmap across algorithms", x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

predicted_ages_corr_heatmap
ggsave("output/figures/predicted_ages_corr_heatmap.jpeg",
  width = 8, height = 6, dpi = 300)
dev.off()
```


```{r}
corr_bd_predicted_ages <- df %>% dplyr::filter(dxgroupFK == 1) %>% dplyr::select(ridge_BrainAge, age_prediction , cnn_BrainAge ,ageprediction_BrainAgeR) %>% cor(.) 

bd_predicted_ages_corr_heatmap <- corr_bd_predicted_ages   %>% melt(.) %>%
  mutate(
    Var1 = factor(Var1, levels = rownames(correlation_matrix), labels = custom_labels),
    Var2 = factor(Var2, levels = rownames(correlation_matrix), labels = custom_labels)
  ) %>%ggplot(., aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "white",   high = "red", midpoint = 0) +
  geom_text(aes(label = sprintf("%.2f", value)), color = "black", size = 4) +
  theme_minimal() +
  labs(title = "Predicted Age correlation heatmap across algorithms in BD", x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))


bd_predicted_ages_corr_heatmap
ggsave("output/figures/bd_predicted_ages_corr_heatmap.jpeg",
  width = 8, height = 6, dpi = 300)
dev.off() 
```


```{r}
# Melt the matrix for ggplot
corr_hc_predicted_ages <- df %>% dplyr::filter(dxgroupFK == 0) %>% dplyr::select(ridge_BrainAge, age_prediction , cnn_BrainAge ,ageprediction_BrainAgeR) %>% cor(.) 

hc_predicted_ages_corr_heatmap <- corr_hc_predicted_ages  %>% melt(.) %>%
  mutate(
    Var1 = factor(Var1, levels = rownames(correlation_matrix), labels = custom_labels),
    Var2 = factor(Var2, levels = rownames(correlation_matrix), labels = custom_labels)
  ) %>% ggplot(., aes(x = Var1, y =Var2, fill=value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "white",   high = "lightblue", midpoint = 0) +
  geom_text(aes(label = sprintf("%.2f", value)), color = "black", size = 4) +
  theme_minimal() +
  labs(title = "Predicted Age correlation heatmap across algorithms in HC", x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))


hc_predicted_ages_corr_heatmap
ggsave("output/figures/hc_predicted_ages_corr_heatmap.jpeg",
  width = 8, height = 6, dpi = 300)
dev.off() 
```

# Standard deviation as a measure of variance in predicted ages
```{r}
df %>% dplyr::select(ridge_BrainAge, age_prediction , cnn_BrainAge ,ageprediction_BrainAgeR, dxgroupFK, gender, agevisit) %>% group_by(dxgroupFK) %>% 
  summarise(sd_predicted_photon = round(sd(age_prediction), 2), 
            sd_predicted_BrainAgeR = round(sd(ageprediction_BrainAgeR), 2), 
            sd_predicted_cnn = round(sd(cnn_BrainAge), 2), 
            sd_predicted_ridge = round(sd(ridge_BrainAge), 2)) 
```


# SCATTER PLOTS SHOWING RELATIONSHIP BETWEEN AGE AND PREDICTED AGE, AGE AND BRAIN-PAD  

### Correlation between age and predicted age for all 4 models; facet by group, and separate lines by sex
```{r }
splot_titles <- c(ridge_BrainAge = 'Multimodal', age_prediction = 'PHOTON', ageprediction_BrainAgeR =  'BrainageR', cnn_BrainAge = 'DenseNet', `0` = 'HC', `1` = 'BD')

df %>% dplyr::select( age_prediction, ageprediction_BrainAgeR , cnn_BrainAge ,ridge_BrainAge, dxgroupFK, gender, agevisit) %>%  melt(id = c('agevisit', 'dxgroupFK', 'gender')) %>%
  ggplot(aes(x = agevisit, y = value , color = factor(gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Predicted Brain Age") +
  theme_minimal() +
  facet_grid(~ variable + dxgroupFK, label = as_labeller(splot_titles)) +  
  geom_text(
    data = . %>% group_by(variable,dxgroupFK) %>% summarise(correlation = cor(agevisit, value, use = "complete.obs" , method = "pearson")),
    aes(x = Inf, y = -Inf, label = paste("r = ", round(correlation, 2)) ),
    hjust = 1, vjust = -1,
    color = "black"
  ) + #s and add annotations
  geom_text(
    data = . %>% group_by(variable,dxgroupFK) %>% summarise(pval = as.numeric(cor.test(agevisit, value)[3])),
    aes(x = Inf, y = -Inf, label =sprintf("p = %.2e", pval)),
    hjust = 1, vjust = 0,
    color = "black"
  ) 
ggsave(filename = "output/figures/predictedage_corr_age.jpeg"  ,dpi = 500, width = 12, height = 3)
dev.off()

```
### Correlation between age and uncorrected brain-PAD for all 4 models, faced by models 
```{r}
splot_titles <- c(bpad_ridge = 'Multimodal', bpad_photon = 'PHOTON', bpad_BrainAgeR =  'BrainageR', bpad_cnn = 'DenseNet',  `0` = 'HC',`1` = 'BD')

df %>% dplyr::select(bpad_photon , bpad_BrainAgeR ,bpad_cnn, bpad_ridge, dxgroupFK, gender, agevisit) %>%  melt(id = c('agevisit', 'dxgroupFK', 'gender')) %>%
  ggplot(aes(x = agevisit, y = value )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Uncorrected Brain-PAD") +
  theme_minimal() +
  facet_grid(~ variable , labeller = as_labeller(splot_titles)  ) + #s and add annotations
  geom_text(
    data = . %>% group_by(variable) %>% summarise(correlation = cor(agevisit, value, use = "complete.obs", method = "pearson")),
    aes(x = Inf, y = -Inf, label = paste("r =", round(correlation, 2)) ),
    hjust = 1, vjust = -1,
    color = "black"
  )  + #s and add annotations
  geom_text(
    data = . %>% group_by(variable) %>% summarise(pval = as.numeric(cor.test(agevisit, value)[3])),
    aes(x = Inf, y = -Inf,  label =sprintf("p = %.2e", pval)),
    hjust = 1, vjust = 0,
    color = "black"
  )   
ggsave("output/figures/uncorrected_bpad_corr_age.jpeg",dpi = 500, width = 12, height = 3)
dev.off()
```

### Correlation between age and uncorrected brain-PAD for all 4 models;facet plots by group, separate lines by sex
```{r fig.align="center", echo = FALSE,fig.width = 12, fig.height=6}
splot_titles <- c(bpad_ridge = 'Multimodal', bpad_photon = 'PHOTON', bpad_BrainAgeR =  'BrainageR', bpad_cnn = 'DenseNet',  `0` = 'HC',  `1` = 'BD')
df %>% dplyr::select(bpad_ridge, bpad_photon , bpad_BrainAgeR ,bpad_cnn, dxgroupFK, gender, agevisit) %>%  melt(id = c('agevisit', 'dxgroupFK', 'gender')) %>%
  ggplot(aes(x = agevisit, y = value , color = factor(gender))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Uncorrected Brain-PAD") +
  theme_minimal() +
  facet_grid(~ variable + dxgroupFK, labeller = as_labeller(splot_titles)) + #s and add annotations
  geom_text(
    data = . %>% group_by(variable,dxgroupFK) %>% summarise(correlation = cor(agevisit, value, use = "complete.obs", method = "pearson")),
    aes(x = Inf, y = -Inf, label = paste("r =", round(correlation, 2)) ),
    hjust= 1, vjust = -1,
    color = "black"
  )  + #s and add annotations
  geom_text(
    data = . %>% group_by(variable,dxgroupFK) %>% summarise(pval = as.numeric(cor.test(agevisit, value)[3])),
    aes(x = Inf, y = -Inf, label =sprintf("p = %.2e", pval)),
    hjust = 1, vjust = 0,
    color = "black"
  )   
ggsave("output/figures/bpad_corr_age_bysexdx.jpeg",dpi = 500, width = 12, height = 3)
dev.off()

```

### Correlation between age and uncorrected brain-PAD for all 4 models;facet plots by group, separate lines by sex
```{r}
splot_titles <- c(ridge_BrainAge = 'Multimodal', age_prediction = 'PHOTON', ageprediction_BrainAgeR =  'BrainageR', cnn_BrainAge = 'DenseNet', `0` = 'HC',  `1` = 'BD')

df %>% dplyr::select(ridge_BrainAge, age_prediction , ageprediction_BrainAgeR , cnn_BrainAge, dxgroupFK, agevisit) %>%  melt(id = c('agevisit', 'dxgroupFK' )) %>%
  ggplot(aes(x = agevisit, y = value )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Corrected Brain-PAD")  +
    scale_color_manual(values=c('black'))+
  theme_minimal() +
  facet_grid(~ variable + dxgroupFK, labeller = as_labeller(splot_titles))    + 
geom_text(data = . %>% group_by(variable, dxgroupFK) %>% summarise(correlation = cor(agevisit, value, use = "complete.obs", method = "pearson")) %>% mutate(correlation = round(correlation, 2)), aes(x = Inf, y = -Inf, label = paste("r =", correlation)), hjust = 1, vjust = -1,  color = "black") + 
geom_text(data = . %>% group_by(variable, dxgroupFK) %>% summarise(pval = as.numeric(cor.test(agevisit, value)[3]))%>% mutate(pval = round(pval, 3)), aes(x = Inf, y = -Inf, label =  paste('p = ',round(pval, 2)),hjust = 1, vjust = 0, color = "black" )) +
  theme(legend.position="none")
ggsave("output/figures/brainage_corr_age_bydx.jpeg",dpi = 500, width = 12, height = 3)
dev.off()
```



### Correlation between age and uncorrected brain-PAD for all 4 models;facet plots by group
```{r}
splot_titles <- c(bpad_ridge = 'Multimodal', bpad_photon = 'PHOTON', bpad_BrainAgeR =  'BrainageR', bpad_cnn = 'DenseNet',  `0` = 'HC',  `1` = 'BD')


df %>% dplyr::select(bpad_ridge, bpad_photon , bpad_BrainAgeR ,bpad_cnn, dxgroupFK, gender, agevisit) %>%  melt(id = c('agevisit', 'dxgroupFK', 'gender')) %>%
  ggplot(aes(x = agevisit, y = value )) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(x = "Chronological Age", y = "Corrected Brain-PAD")  +
    scale_color_manual(values=c('black'))+
  theme_minimal() +
  facet_grid(~ variable + dxgroupFK, labeller = as_labeller(splot_titles))    + 
geom_text(data = . %>% group_by(variable, dxgroupFK) %>% summarise(correlation = cor(agevisit, value, use = "complete.obs", method = "pearson")) %>% mutate(correlation = round(correlation, 2)), aes(x = Inf, y = -Inf, label = paste("r =", correlation)), hjust = 1, vjust = -1,  color = "black") + 
geom_text(data = . %>% group_by(variable, dxgroupFK) %>% summarise(pval = as.numeric(cor.test(agevisit, value)[3]))%>% mutate(pval = round(pval, 3)), aes(x = Inf, y = -Inf, label =  paste('p = ',round(pval, 2)),hjust = 1, vjust = 0, color = "black" )) +
  theme(legend.position="none")
ggsave("output/figures/bpad_corr_age_bydx.jpeg",dpi = 500, width = 12, height = 3)
dev.off()
```



