---
title: "Demographics"
author: "Renee Hui Xin Ng"
date: "2025-04-17"
output: html_document
---

# Load required libraries
```{r setup, include=FALSE}
library(dplyr)
library(tidyr)       # used with dplyr pipelines
library(ggplot2)
library(reshape2)    # for melt() – even though deprecated, you're still using it
library(readxl)      # if reading Excel files, although not showNn in session
library(haven)       # for `read_sav()` when loading SPSS files
library(DescTools)   # for `icc()` function
library(data.table)  # used with `setDT()` and `melt()`
library(lme4)        # for `lmer()`
library(lmerTest)    # for p-values in `lmer`
library(lmtest)      # optional, used for model testing
#library(irr)         # for inter-rater reliability (e.g. `icc()`, alternative to DescTools)
library(emmeans)
```

# What does this rmd file accomplish?
Visualization: 
- Density plots to assess the distribution of Brain-PAD values and predicted brain ages across models
- Scatter plots to explore the relationship between chronological age and brain age predictions.

Normality Testing: 
Shapiro-Wilk tests to evaluate the normality of Brain-PAD and predicted age distributions across different models and groups.

Correlation Analysis:
Correlations between chronological age and Brain-PAD values, as well as between predicted ages across models, including separate analyses for the HC and BD groups. The results are presented in correlation matrices and heatmaps.

Standard dev: SD of Brain-PAD and predicted age values as measures of variance within both HC and BD groups.
 
# Load and clean original dataset (Master_BAI.....) but we may not use it
```{r original df}
# this is ocassionally used for sanity checks
path = file.path("../data_original/Master_BAI_080720_with_051517completers_060117imagingmaster_genotypes_brainages_ENIGMA_FS_MERGEFK_30122020.sav")
original_df = read_sav(path)

original_df <- original_df %>% dplyr::select("trnsctrw", "cwiisnrs","cwiisis", "cwiicns","fsrp5","stdyptid", 'visit', 'dxgroupFK', 'agevisit', 'HistoryPsychosis', Log_DepressiveEpisodes, NumManicEpisodes_Log, AgeofOnset, psytrtot, fsrptot , scid_a1, scid_h5, npsyttot, 'hamtot17', 'hamtot28', 'ymrstot', psytr1, CurrentMedLoad, AgeofOnset, veteran, educ_cat, educ, hisp, livsit) %>% filter(!is.na(trnsctrw)|!is.na(fsrp5)) %>% arrange(stdyptid) 



original_df

original_df$timeSinceOnset <- original_df$agevisit- original_df$AgeofOnset

# dosage they were taking were rated -
```

# full dataset with all brain ages
```{r full dataset from collate_all_brainages.ipynb}
df<-read.csv("../data_processed/df_all_brainage_2025.csv") 
df %>% group_by(dxgroupFK) %>%tally()
df  <- df %>% mutate(Age2 = agevisit**2, Age3 = agevisit**3, 
                     bpad_ridge = ridge_BrainAge - agevisit,
                     bpad_photon = age_prediction - agevisit,
                     bpad_BrainAgeR = ageprediction_BrainAgeR - agevisit,
                     bpad_cnn = cnn_BrainAge - agevisit)

merged_df <- merge(df, original_df, by = c('stdyptid' , 'dxgroupFK')) # %>% dplyr::select('hamtot17.y',  dxgroupFK) #%>% filter(!is.na(hamtot17.y))


df[, grepl("Current" , names( df ))]

df %>% group_by(dxgroupFK) %>% tally() 
```

# Create derived variables including Brain-PAD (predicted age - chronological age)
```{r}
summary_df <- df  %>% filter(
    !is.na(ridge_BrainAge), !is.na(age_prediction), !is.na(ageprediction_BrainAgeR), !is.na(cnn_BrainAge),
    !is.na(bpad_ridge), !is.na(bpad_photon), !is.na(bpad_BrainAgeR), !is.na(bpad_cnn)
  ) %>%
  summarise(
    # Predicted Brain Ages
    ridge_BrainAge = paste0(round(mean(ridge_BrainAge, na.rm = TRUE), 1), " (", round(sd(ridge_BrainAge, na.rm = TRUE), 2), ")"),
    age_prediction = paste0(round(mean(age_prediction, na.rm = TRUE), 1), " (", round(sd(age_prediction, na.rm = TRUE), 2), ")"),
    ageprediction_BrainAgeR = paste0(round(mean(ageprediction_BrainAgeR, na.rm = TRUE), 1), " (", round(sd(ageprediction_BrainAgeR, na.rm = TRUE), 2), ")"),
    cnn_BrainAge = paste0(round(mean(cnn_BrainAge, na.rm = TRUE), 1), " (", round(sd(cnn_BrainAge, na.rm = TRUE), 2), ")"),
    
    # Brain-PADs
    bpad_ridge = paste0(round(mean(bpad_ridge, na.rm = TRUE), 2), " (", round(sd(bpad_ridge, na.rm = TRUE), 2), ")"),
    bpad_photon = paste0(round(mean(bpad_photon, na.rm = TRUE), 2), " (", round(sd(bpad_photon, na.rm = TRUE), 2), ")"),
    bpad_BrainAgeR = paste0(round(mean(bpad_BrainAgeR, na.rm = TRUE), 2), " (", round(sd(bpad_BrainAgeR, na.rm = TRUE), 2), ")"),
    bpad_cnn = paste0(round(mean(bpad_cnn, na.rm = TRUE), 2), " (", round(sd(bpad_cnn, na.rm = TRUE), 2), ")")
  )
 

```
 
# HC = 73 Demographics
```{r}
test_full_hc_df <- df %>%
  group_by(dxgroupFK)  
# Overall test for age differences across groups
age_test <- aov(agevisit ~ dxgroupFK, data = test_full_hc_df)
summary(age_test)


sex_summary <- test_full_hc_df %>%
  group_by(dxgroupFK, gender) %>%
  summarise(N = n()) %>%
  mutate(percent = round(N / sum(N) * 100, 1)) %>%
  arrange(dxgroupFK, desc(gender))
sex_summary

# Overall test for sex differences across groups
sex_test <- chisq.test(table(test_full_hc_df$dxgroupFK, test_full_hc_df$gender))
print(sex_test)

# Example: summarize age
age_summary <- test_full_hc_df %>% 
  summarise(
    Mean_Age = round(mean(agevisit, na.rm = TRUE), 1),
    SD_Age = round(sd(agevisit, na.rm = TRUE), 2),
    N = n()
  )
age_summary

 
test_full_hc_df

 
# List of variables you want to run ANOVA on
variables_to_test <- c("agevisit", "age_prediction", "ageprediction_BrainAgeR", 
                       "cnn_BrainAge", "bpad_photon", "bpad_BrainAgeR", 'bpad_cnn')

# Run ANOVA for each variable
anova_results <- lapply(variables_to_test, function(var) {
  formula <- as.formula(paste(var, "~ dxgroupFK"))
  aov(formula, data = test_full_hc_df)
})

# Name the list so you know which result is which
names(anova_results) <- variables_to_test

# To view the summary of each ANOVA result
lapply(anova_results, summary)


# Example: summarize age
test_full_hc_df %>% 
  summarise(
    Mean_Age = round(mean(bpad_cnn, na.rm = TRUE), 1),
    SD_Age = round(sd(bpad_cnn, na.rm = TRUE), 2),
    N = n()
  )





```

### HC = 22 Demographics data
```{r}
test_df <- df %>%
  group_by(dxgroupFK)%>% filter(!is.na(ridge_BrainAge))
# Overall test for age differences across groups
age_test <- aov(agevisit ~ dxgroupFK, data = test_df)
summary(age_test)

 # Example: summarize sex
sex_summary <- test_df %>%
  group_by(dxgroupFK, gender) %>%
  summarise(N = n()) %>%
  mutate(percent = round(N / sum(N) * 100, 1)) %>%
  arrange(dxgroupFK, desc(gender))

# Example: summarize age
age_summary <- test_df%>%
  summarise(
    Mean_Age = round(mean(agevisit, na.rm = TRUE), 1),
    SD_Age = round(sd(agevisit, na.rm = TRUE), 2),
    N = n())
age_summary

# Overall test for sex differences across groups
sex_test <- chisq.test(table(test_df$dxgroupFK, test_df$gender))
print(sex_test)
 
# List of variables you want to run ANOVA on
variables_to_test <- c("agevisit", "ridge_BrainAge", "age_prediction", "ageprediction_BrainAgeR", 
                       "cnn_BrainAge", "bpad_ridge", "bpad_photon", "bpad_BrainAgeR", 'bpad_cnn')

# Run ANOVA for each variable
anova_results <- lapply(variables_to_test, function(var) {
  formula <- as.formula(paste(var, "~ dxgroupFK"))
  aov(formula, data = test_df)
})

# Name the list so you know which result is which
names(anova_results) <- variables_to_test

# To view the summary of each ANOVA result
lapply(anova_results, summary)

temp




threeway_photon_bpad<-ggplot(temp, aes(x = agevisit.y, y = bpad_photon, color = as.factor(dxgroupFK))) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = TRUE, size = 1.2) +
  scale_color_manual(
    name = "Group",
    values = c("1" = "red", "0" = "blue"),
    labels = c("1" = "BD", "0" = "HC")
  ) +
  coord_cartesian(ylim = c(-40, 40)) +
  facet_wrap(~ gender) +  # Facet by diagnosis group
  labs(
    x = "Chronological Age",
    y = "PHOTON Brain-PAD",
    caption = "Linear fit with 95% CI"
  ) +
  theme_minimal(base_size = 14)

ggsave("output/figures/three_way_interaction_photon_bpad.jpeg",dpi = 500, width = 10, height = 5)
threeway_photon_bpad
dev.off()
temp %>% group_by(dxgroupFK, gender) %>% tally()

```

# create temp df to run linear regression models related to clinical factors
- HistoryPsychosis.y (bpad_cnn is significant)
- Log_DepressiveEpisodes.y
- NumManicEpisodes_Log.y

# Lithium vs HC

```{r}
temp <- merged_df %>% dplyr::select( c("agevisit.y", "ridge_BrainAge", "age_prediction", "ageprediction_BrainAgeR", 
                       "cnn_BrainAge" ,"stdyptid", 'dxgroupFK',   'hamtot17.y','hamtot28.y','gender', 'visit.y',HistoryPsychosis.y, NumManicEpisodes_Log.y, Log_DepressiveEpisodes.y, 'AntiEpileptic', 'Gen2AntiPsych', 'Gen1AntiPsych', 'AntiDep', 'Li', 'CurrentMedLoad.y', AgeofOnset.y, psytrtot.y,  'scid_a1.y', 'scid_h5.y', 'npsyttot.y', timeSinceOnset))  %>% 
            mutate(Age2 = agevisit.y**2, Age3 = agevisit.y**3, 
                     bpad_ridge = ridge_BrainAge - agevisit.y,
                     bpad_photon = age_prediction - agevisit.y,
                     bpad_BrainAgeR = ageprediction_BrainAgeR - agevisit.y,
                     bpad_cnn = cnn_BrainAge - agevisit.y) %>% filter(visit.y == 0)


merged_df
temp_lithium <- temp %>%
  mutate(lithium_groups = case_when(
    dxgroupFK == 0 ~ "HC",
    dxgroupFK == 1 & Li == 1 ~ "BD_Li",
    dxgroupFK == 1 & Li == 0~ "BD_No_Li",
    TRUE ~ NA_character_
  )) %>%
  filter(!is.na(lithium_groups))

temp_aed <- temp %>%
  mutate(aed_groups = case_when(
    dxgroupFK == 0 ~ "HC",
    dxgroupFK == 1 & AntiEpileptic == 1 ~ "BD_AED",
    dxgroupFK == 1 & AntiEpileptic == 0~ "BD_No_AED",
    TRUE ~ NA_character_
  )) %>%
  filter(!is.na(aed_groups))

temp_aed %>% group_by(aed_groups) %>% tally()
temp_lithium %>% group_by(lithium_groups) %>% tally()

temp_aed

res.aov <- aov(agevisit.y ~ aed_groups, data = temp_aed)
summary(res.aov)

temp_aed

### AED

tbl <-table(temp_aed$aed_groups, as.character(temp_aed$HistoryPsychosis.y))
chisq.test(tbl)

# Post-hoc ANOVA analyses revealed?

res.aov <- aov(CurrentMedLoad.y~ aed_groups, data = temp_aed)
summary(res.aov) # sig


temp_aed %>%
  group_by(aed_groups) %>%
  summarise(
    mean_medload = mean(CurrentMedLoad.y, na.rm = TRUE),
    sd_medload = sd(CurrentMedLoad.y, na.rm = TRUE),
    n = n()
  )
 
res.aov <- aov(psytrtot.y ~ aed_groups, data = temp_aed)
summary(res.aov) # sig


temp_aed %>%
  group_by(aed_groups) %>%
  summarise(
    mean_medload = mean(psytrtot.y, na.rm = TRUE),
    sd_medload = sd(psytrtot.y, na.rm = TRUE),
    n = n()
  )

# is there something else 


res.aov <- aov(HistoryPsychosis.y ~ aed_groups, data = temp_aed)
summary(res.aov)


res.aov <- aov(Log_DepressiveEpisodes.y ~ aed_groups, data = temp_aed)
summary(res.aov)


res.aov <- aov(NumManicEpisodes_Log.y ~ aed_groups, data = temp_aed)
summary(res.aov)
  

res.aov <- aov(timeSinceOnset ~ aed_groups, data = temp_aed)
summary(res.aov)



# recode reference group
temp_lithium$lithium_groups <- relevel(factor(temp_lithium$lithium_groups), ref = "HC")
temp_aed$aed_groups <- relevel(factor(temp_aed$aed_groups), ref = "HC")
```

# BD li vs BD no li vs HC
```{r}
mod <- lm(bpad_photon ~ lithium_groups + agevisit.y + gender, data = temp_lithium)
 
summary(mod)
anova(mod) # sig


# PHOTON
em_photon_lithium <- emmeans(mod, pairwise ~ lithium_groups)
summary(em_photon_lithium)

mod <- lm(bpad_ridge ~ lithium_groups + agevisit.y + gender, data = temp_lithium)
summary(mod)
anova(mod) # trend

mod <- lm(bpad_cnn ~ lithium_groups + agevisit.y + gender, data = temp_lithium)
summary(mod)
anova(mod) # trend

mod <- lm(bpad_BrainAgeR ~ lithium_groups + gender , data = temp_lithium)
summary(mod)
anova(mod) # none
```

# BD AED vs BD no AED vs HC
```{r}
mod <- lm(bpad_photon ~ aed_groups + agevisit.y + gender, data = temp_aed)
summary(mod)
anova(mod) # sig

# PHOTON
em_photon_aed <- emmeans(mod, pairwise ~ aed_groups)
summary(em_photon_aed)

mod <- lm(bpad_cnn ~ aed_groups + agevisit.y + gender, data = temp_aed)
summary(mod)
anova(mod) # sig

# CNN
em_cnn_aed <- emmeans(mod, pairwise ~ aed_groups)
summary(em_cnn_aed)
 
mod <- lm(bpad_ridge ~ aed_groups + agevisit.y + gender, data = temp_aed)
summary(mod)
anova(mod)

mod <- lm(bpad_BrainAgeR ~ aed_groups + gender , data = temp_aed)
summary(mod)
anova(mod)


# Define a custom plotting function
plot_violin <- function(data, group_var, pad_var, model_label) {
  ggplot(data, aes_string(x = group_var, y = pad_var, fill = group_var)) +
    geom_violin(trim = FALSE, alpha = 0.6) +
    geom_boxplot(width = 0.1, outlier.shape = NA) +
    stat_summary(fun = mean, geom = "point", shape = 20, size = 2, color = "black") +
    labs(title = paste(model_label,"Brain-PAD", "across BD medication groups () vs HC"),
         x = "", y = "Brain-PAD") +
    coord_cartesian(ylim = c(-40, 40)) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "none")
}




violin_plot_data <- temp %>%# filter(dxgroupFK == 1) %>%
  dplyr::select(dxgroupFK, stdyptid,bpad_photon, bpad_BrainAgeR, bpad_cnn, bpad_ridge) %>%
  pivot_longer(cols = starts_with("bpad_"),
               names_to = "model",
               values_to = "brain_PAD") %>%
  mutate(model = recode(model,
                        bpad_photon = "PHOTON",
                        bpad_BrainAgeR = "BrainAgeR",
                        bpad_cnn = "DenseNet",
                        bpad_ridge = "Multimodal"))

violin_plot_data

summary_stats_bd <- violin_plot_data %>%filter(dxgroupFK == 1) %>%
  group_by(model) %>%
  summarise(
    mean_val = mean(brain_PAD, na.rm = TRUE),
    sd_val = sd(brain_PAD, na.rm = TRUE)
  )



violin_plot_data_bd <- violin_plot_data  %>%filter(dxgroupFK == 1)

# Set factor levels in the desired order
violin_plot_data_bd$model <- factor(violin_plot_data_bd$model, levels = c("PHOTON", "BrainAgeR", "DenseNet", "Multimodal"))

summary_stats_bd$model <- factor(summary_stats_bd$model, levels = c("PHOTON", "BrainAgeR", "DenseNet", "Multimodal"))

violin_plot_data_bd%>% ggplot(., aes(x = model, y = brain_PAD, fill = 'purple')) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 2.5, color = "black") +
  labs(title = "Brain-PAD across models in BD participants",
       x = "Model", y = "Brain-PAD") +
  coord_cartesian(ylim = c(-40, 40)) +
  theme_minimal(base_size = 14) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 2.5, color = "black") +
  geom_text(data = summary_stats_bd,
            aes(x = model, y = 35,  # Adjust Y-position as needed
                label = paste0("Mean = ", round(mean_val, 2), 
                               "\nSD = ", round(sd_val, 2))),
            size = 4, color = "black", inherit.aes = FALSE) +
  labs(#title = "Brain-PAD across models in BD participants",
       x = "Model", y = "Brain-PAD") +
  coord_cartesian(ylim = c(-40, 40)) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")




violin_plot_data
# Set factor levels in the desired order
violin_plot_data$model <- factor(violin_plot_data$model, levels = c("PHOTON", "BrainAgeR", "DenseNet", "Multimodal"))
summary_stats$model <- factor(summary_stats$model, levels = c("PHOTON", "BrainAgeR", "DenseNet", "Multimodal"))


violin_plot_data$dxgroupFK <- factor(violin_plot_data$dxgroupFK,
                                     levels = c(0, 1),
                                     labels = c("HC", "BD"))
summary_stats$dxgroupFK <- factor(summary_stats$dxgroupFK,
                                  levels = c(0, 1),
                                  labels = c("HC", "BD"))

summary_stats <- violin_plot_data %>%
  group_by(model, dxgroupFK) %>%
  summarise(
    mean_val = mean(brain_PAD, na.rm = TRUE),
    sd_val = sd(brain_PAD, na.rm = TRUE),
    .groups = "drop"
  )
# Plot

ggplot(violin_plot_data, aes(x = model, y = brain_PAD, fill = as.factor(dxgroupFK))) +
  geom_violin(trim = FALSE, alpha = 0.6, position = position_dodge(width = 0.8)) +
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7,
               position = position_dodge(width = 0.8)) +
  stat_summary(fun = mean, geom = "point", shape = 20, size = 2.5, color = "black",
               position = position_dodge(width = 0.8)) +
  geom_text(data = summary_stats,
            aes(x = model,
                y = 39,  # Move further up
                group = dxgroupFK,
                label = paste0("μ= ", round(mean_val, 1),
                               "\nSD= ", round(sd_val, 1))),
            position = position_dodge(width = 0.8),
            size = 3.5,
            color = "black",
            inherit.aes = FALSE) +
  labs(
       x = "Algorithm", y = "Brain-PAD", fill = "Group") +
  scale_fill_manual(values = c("BD" = "red", "HC" = "lightblue")) +
  coord_cartesian(ylim = c(-40, 40)) +
  theme_minimal(base_size = 14)



# RAW Brain-PADs
# Lithium group plots
ggsave("violin_PHOTON_lithium.jpeg", plot_violin(temp_lithium, "lithium_groups", "bpad_photon", "PHOTON"), width = 8, height = 4, dpi = 300)
ggsave("violin_Ridge_lithium.jpeg", plot_violin(temp_lithium, "lithium_groups", "bpad_ridge", "Multimodal Ridge"), width = 8, height = 4, dpi = 300)
ggsave("violin_DenseNet_lithium.jpeg", plot_violin(temp_lithium, "lithium_groups", "bpad_cnn", "DenseNet"), width = 8, height = 4, dpi = 300)
ggsave("violin_BrainageR_lithium.jpeg", plot_violin(temp_lithium, "lithium_groups", "bpad_BrainAgeR", "BrainageR"), width = 8, height = 4, dpi = 300)

# AED group plots
ggsave("violin_PHOTON_aed.jpeg", plot_violin(temp_aed, "aed_groups", "bpad_photon", "PHOTON"), width = 8, height = 4, dpi = 300)
ggsave("violin_Ridge_aed.jpeg", plot_violin(temp_aed, "aed_groups", "bpad_ridge", "Multimodal Ridge"), width = 8, height = 4, dpi = 300)
ggsave("violin_DenseNet_aed.jpeg", plot_violin(temp_aed, "aed_groups", "bpad_cnn", "DenseNet"), width = 8, height = 4, dpi = 300)
ggsave("violin_BrainageR_aed.jpeg", plot_violin(temp_aed, "aed_groups", "bpad_BrainAgeR", "BrainageR"), width = 8, height = 4, dpi = 300)

# small sample of people on lithium were indistinguishable from HC
# neither were people who were not on lithium

plot_violin(temp_aed, "aed_groups", "bpad_ridge", "PHOTON")
```

# We only plot emmeans for the models that have significant omnibust test 

```{r}

plot_emmeans_with_ci <- function(emm_grid, group_var, title, filename,
                                 width = 8, height = 5, dpi = 300) {
  # Convert to data frame
  em_df <- as.data.frame(emm_grid)

  # Ensure group_var is a symbol
  group_var_sym <- rlang::sym(group_var)

  # Plot with error bars
  p <- ggplot(em_df, aes(x = !!group_var_sym, y = emmean)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
    labs(
      title = title,
      y = "Estimated Brain-PAD",
      x = group_var
    ) +
    theme_minimal(base_size = 14) + coord_flip()

  # Save plot
  ggsave(filename, plot = p, width = width, height = height, dpi = dpi)
}

# AED 
em_df <- as.data.frame(em_photon_aed$emmeans)
print(em_df)
plot_emmeans_with_ci(
  emm_grid = em_df,
  group_var = "aed_groups",
  title = "PHOTON EMM of brain-PAD across AED vs non-AED vs HC",
  filename = "emmean_PHOTON_AED.jpeg"
)

em_df <- as.data.frame(em_cnn_aed$emmeans)
print(em_df)

plot_emmeans_with_ci(
  emm_grid = em_df,
  group_var = "aed_groups",
  title = "DenseNet EMM of brain-PAD across AED vs non-AED vs HC",
  filename = "emmean_cnn_AED.jpeg"
)


em_df <- as.data.frame(em_photon_lithium$emmeans)
print(em_df)

plot_emmeans_with_ci(
  emm_grid = em_df,
  group_var = "lithium_groups",
  title = "PhOTHON EMM of brain-PAD across Li vs non-Li vs HC",
  filename = "emmean_photon_lithium.jpeg"
)
```

```{r}
mod <- lm(bpad_BrainAgeR ~ gender  + timeSinceOnset , data = temp)
summary(mod)

mod <- lm(bpad_photon ~ gender + agevisit.y + timeSinceOnset , data = temp)
summary(mod)


mod <- lm(bpad_ridge ~ gender +agevisit.y + timeSinceOnset , data = temp)
summary(mod)

mod <- lm(bpad_cnn ~ gender +agevisit.y + timeSinceOnset , data = temp)
summary(mod)
```

```{r}
# Compare BD NOT on lithium vs. HCs

# temp_no_li <-temp %>% filter(is.na(Li) | Li == 0)
# 
# mod_photon <- lm(bpad_photon ~  dxgroupFK+ agevisit.y , data = temp_no_li)
# summary(mod_photon)
# 
# mod_ridge <- lm(bpad_ridge ~  dxgroupFK+ agevisit.y , data = temp_no_li)
# summary(mod_ridge)
# 
# mod_cnn <- lm(bpad_cnn ~  dxgroupFK+ agevisit.y , data = temp_no_li)
# summary(mod_cnn)
# 
# mod_BrainAgeR<- lm(bpad_BrainAgeR ~  dxgroupFK+ agevisit.y , data = temp_no_li)
# summary(mod_BrainAgeR)
```

```{r}
# Compare BD on lithium vs. HCs
# temp_using_li <-temp %>% filter(is.na(Li) | Li == 1)
# 
# temp_using_li
# mod_photon <- lm(bpad_photon ~  dxgroupFK+ agevisit.y , data = temp_using_li)
# summary(mod_photon)
# 
# mod_ridge <- lm(bpad_ridge ~  dxgroupFK+ agevisit.y , data = temp_using_li)
# summary(mod_ridge)
# 
# mod_cnn <- lm(bpad_cnn ~  dxgroupFK+ agevisit.y , data = temp_using_li)
# summary(mod_cnn)
# 
# mod_BrainAgeR<- lm(bpad_BrainAgeR ~  dxgroupFK+ agevisit.y , data = temp_using_li)
# summary(mod_BrainAgeR)
```


```{r}
# 1. Subset to BD only
bd <- temp %>% filter(dxgroupFK == 1)

bd$scid_a1
# 2. Define your variables
cat_vars  <- c("HistoryPsychosis.y", "Li", "AntiEpileptic",
               "Gen2AntiPsych", "Gen1AntiPsych", "AntiDep")
cont_vars <- c("AgeofOnset.y", "psytrtot.y", 
               "NumManicEpisodes_Log.y", "Log_DepressiveEpisodes.y",  'scid_a1.y', 'scid_h5.y', 'npsyttot.y', 'CurrentMedLoad.y')

# 3. A little lookup for nicer names
nice <- c(
  HistoryPsychosis.y           = "History of Psychosis",
  Li                            = "Lithium",
  AntiEpileptic                 = "AED",
  Gen2AntiPsych                 = "SGA",
  Gen1AntiPsych                 = "FGA",
  AntiDep                       = "Antidepressants",
  AgeofOnset.y                  = "Age of Onset",
  psytrtot.y                    = "Total Psyc Medications",
  NumManicEpisodes_Log.y        = "N Manic Episodes (Log)",
  Log_DepressiveEpisodes.y      = "N Depressive Episodes (Log)",
  CurrentMedLoad.y = 'MedLoad',
  scid_a1.y = 'Current GAF (SCID)',
  scid_h5.y = 'N Suicide Attempts', 
  npsyttot.y = 'Total Non-Psyc Medication'
  
)

library(purrr)
# 4. Summarise categoricals
cat_summary <- map_dfr(cat_vars, function(var) {
  n   <- sum(bd[[var]] == 1, na.rm = TRUE)
  avl <- sum(!is.na(bd[[var]]))
  tibble(
    Characteristic = nice[var],
    N              = n,
    Available      = avl,
    Mean           = NA_real_,
    SD             = NA_real_,
    Percent        = round(100 * n / avl, 1)
  )
})

# 5. Summarise continuous
cont_summary <- map_dfr(cont_vars, function(var) {
  avl  <- sum(!is.na(bd[[var]]))
  m    <- round(median(bd[[var]], na.rm = TRUE), 2)
  s    <- round(IQR(bd[[var]],   na.rm = TRUE), 2)
  tibble(
    Characteristic = nice[var],
    N              = NA_integer_,
    Available      = avl,
    Mean           = m,
    SD             = s,
    Percent        = NA_real_
  )
})

# 6. Combine
summary_df <- bind_rows(cat_summary, cont_summary)

# 7. Re-order if you like
summary_df <- summary_df %>% arrange(match(Characteristic, nice))

print(summary_df)

```

# HistoryPsychosis ~ bpad + age
Other tested: Number of Manic Episodes, Hamilton 28, YMRstotal 
```{r}
mod <- lm( bpad_photon~  HistoryPsychosis.y  + agevisit.y + gender, data = temp)
summary(mod)

mod <- lm( bpad_ridge  ~ HistoryPsychosis.y  + agevisit.y+ gender, data = temp)
summary(mod)

mod <- lm( bpad_BrainAgeR~  HistoryPsychosis.y + gender, data = temp)
summary(mod)

mod <- lm(bpad_cnn ~  HistoryPsychosis.y+ agevisit.y+ gender, data = temp)
summary(mod)
```


#NumManicEpisodes_Log.y 
```{r}
mod <- glm(bpad_photon  ~ NumManicEpisodes_Log.y  + agevisit.y + gender, data = temp)
summary(mod)

mod <- glm( bpad_ridge ~  NumManicEpisodes_Log.y+ agevisit.y+ gender, data = temp)
summary(mod)

mod <- glm(bpad_BrainAgeR  ~ NumManicEpisodes_Log.y + gender , data = temp)
summary(mod)

mod <- glm(bpad_cnn  ~ NumManicEpisodes_Log.y + agevisit.y+ gender, data = temp)
summary(mod)
```
# NumManicEpisodes_Log.y 
```{r}
mod <- glm( bpad_photon ~   CurrentMedLoad.y+ agevisit.y + gender, data = temp)
summary(mod)

mod <- glm( bpad_ridge ~  CurrentMedLoad.y + agevisit.y+ gender, data = temp)
summary(mod)

mod <- glm( bpad_BrainAgeR ~CurrentMedLoad.y  + gender , data = temp)
summary(mod)

mod <- glm( bpad_cnn ~   CurrentMedLoad.y+ agevisit.y+ gender, data = temp)
summary(mod)



#NumManicEpisodes_Log.y 

```

 
# Group differences divided by medication
```{r}
temp <- temp %>%
  mutate(moodStabilizer = 
                  case_when(
                    Li == 0 & AntiEpileptic == 0~ 'None', 
                    Li == 1  & AntiEpileptic == 0~ 'Li', #  Any Li
                    AntiEpileptic == 1 &   Li == 0 ~ 'AED', # Any AED
                  Li == 1 & AntiEpileptic == 1 ~ 'AED + Li')) # On AED + Li

med_count<-as.data.table(temp) %>% group_by(moodStabilizer) %>% tally() %>% na.omit() #%>% kable() # how many data points in each group?
med_count

temp
summary_tbl <- temp %>%
  group_by(moodStabilizer) %>%
  summarise(
    N           = n_distinct(stdyptid),
    male_n      = sum(gender == 1, na.rm = TRUE),
    male_pct    = round(100 * male_n / N, 1),
    age_mean    = round(mean(agevisit.y, na.rm = TRUE), 2),
    age_sd      = round(sd(agevisit.y,   na.rm = TRUE), 2),
    NumPsychotropic = round(mean(psytrtot.y, na.rm = TRUE), 2),
    medload = round(median(CurrentMedLoad.y, na.rm = TRUE), 2),
    # now all four bpads:
    across(
      .cols = starts_with("bpad_"),
      .fns  = list(mean = ~ round(mean(.x, na.rm = TRUE), 2),
                   sd   = ~ round(sd(.x,   na.rm = TRUE), 2)),
      .names = "{.col}_{.fn}"
    ), .groups = "drop"
  )

print(summary_tbl)

# Table 1 stats
tbl <-table(temp$moodStabilizer, as.character(temp$gender))
chisq.test(tbl)

res.aov <- aov(agevisit.y ~ moodStabilizer, data = temp)
summary(res.aov)



tbl <-table(temp$moodStabilizer, as.character(temp$HistoryPsychosis.y))
chisq.test(tbl)

# Post-hoc ANOVA analyses revealed that medication load (p < 0.001) and number of antipsychotic medications (p < 0.01)   were significantly different across the medication groups of AED (n = 18) and lithium monotherapy (n = 5), combination of the two (n = 4)
res.aov <- aov(CurrentMedLoad.y~ moodStabilizer, data = temp)
summary(res.aov)

 
res.aov <- aov(psytrtot.y ~ moodStabilizer, data = temp)
summary(res.aov)


res.aov <- aov(CurrentMedLoad.y ~ moodStabilizer, data = temp)
summary(res.aov)


res.aov <- aov(Log_DepressiveEpisodes.y ~ moodStabilizer, data = temp)
summary(res.aov)


res.aov <- aov(NumManicEpisodes_Log.y ~ moodStabilizer, data = temp)
summary(res.aov)


# NumManicEpisodes_Log

temp %>% group_by(moodStabilizer)  %>%
  summarise(n = n())

library(broom)
# temp %>% 
#   select(starts_with("bpad_"), moodStabilizer) %>%
#   pivot_longer(-moodStabilizer, names_to = "measure", values_to = "value") %>%
#   group_by(measure) %>%
#   do(tidy(aov(value ~ moodStabilizer, data = .))) %>%
#   filter(term == "moodStabilizer") %>%
#   select(measure, statistic, p.value) 
```

# 
```{r}
moodMod_photon <- lm(bpad_photon ~ agevisit.y + gender + Log_DepressiveEpisodes.y , data = temp)
summary(moodMod_photon) # sig




moodMod_brainageR <- lm(bpad_BrainAgeR ~ agevisit.y + gender +Log_DepressiveEpisodes.y  , data = temp)
summary(moodMod_brainageR)


moodMod_cnn <- lm(bpad_cnn ~ agevisit.y + gender + Log_DepressiveEpisodes.y , data = temp)
summary(moodMod_cnn) 


moodMod_ridge <- lm(bpad_ridge ~ agevisit.y + gender + Log_DepressiveEpisodes.y , data = temp)
summary(moodMod_ridge)

```

# AED * Li
```{r}
moodMod_photon <- lm(bpad_photon ~ agevisit.y + gender + AntiEpileptic*Li , data = temp)
summary(moodMod_photon)

# Add predicted values

temp_bd$predicted_brain_pad_li_aed_interaction <- predict(moodMod_photon)

# Create a group variable for interaction
temp_bd <- temp %>%
  filter(dxgroupFK == 1) %>%
  mutate(
    predicted_brain_pad_li_aed_interaction = predict(moodMod_photon),
    group = interaction(AntiEpileptic, Li, sep = "_")
  )

# Label the groups for clarity
temp_bd$group <- recode(temp_bd$group,
                        `0_0` = "No AED, No Li",
                        `1_0` = "AED",
                        `0_1` = "Li",
                        `1_1` = "AED + Li")

# Plot predicted Brain-PAD by group
ggplot(temp_bd, aes(x = group, y = predicted_brain_pad_li_aed_interaction)) +
  #geom_point(position = position_jitter(width = 0.1), alpha = 0.6) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 3) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2) +
  labs(x = "Medication Group", y = "Predicted Brain-PAD",
       title = "Interaction of AED and Lithium on Predicted Brain-PAD") +
  theme_minimal(base_size = 14)


```

```{r}

# Get EMMeans for each combination of AntiEpileptic and Li
em <- emmeans(moodMod_photon, ~ AntiEpileptic * Li)
em
em_df<- as.data.frame(em)
print(em_df)
# Optional: label the interaction groups for readability
em_df <- em_df %>%
  mutate(group = interaction(AntiEpileptic, Li, sep = "_")) %>%
  mutate(group = recode(group,
                        `0_0` = "No AED, No Li",
                        `1_0` = "AED",
                        `0_1` = "Li",
                        `1_1` = "AED + Li"))
em_df
# Plot
ggplot(em_df, aes(x = group, y = emmean)) +
  geom_point(color = "black", size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  labs(
    x = "Medication",
    y = "Estimated Brain-PAD",
    #title = 
  ) +
  theme_minimal(base_size = 14) + coord_flip()


moodMod_brainageR <- lm(bpad_BrainAgeR~  gender + AntiEpileptic*Li, data = temp)
summary(moodMod_brainageR)

# Coefficients:
#                  Estimate Std. Error t value Pr(>|t|)  
# (Intercept)        2.6368     2.4609   1.071   0.2915  
# gender             3.2950     2.7372   1.204   0.2370  
# AntiEpileptic     -4.9945     2.8972  -1.724   0.0938 .
# Li                 0.5332     4.0212   0.133   0.8953  
# AntiEpileptic:Li   5.0303     5.8756   0.856   0.3979  
moodMod_cnn <- lm(bpad_cnn~ agevisit.y + gender + AntiEpileptic*Li , data = temp)
summary(moodMod_cnn)

# 
# Coefficients:
#                  Estimate Std. Error t value Pr(>|t|)    
# (Intercept)        4.4731     4.0721   1.098 0.279951    
# agevisit.y        -0.1571     0.0788  -1.994 0.054523 .  
# gender             1.4457     1.6908   0.855 0.398692    
# AntiEpileptic     -6.9797     1.7888  -3.902 0.000444 ***
# Li                 1.1012     2.5187   0.437 0.664808    
# AntiEpileptic:Li   1.2156     3.6339   0.335 0.740099  

moodMod_ridge<- lm(bpad_ridge~ agevisit.y + gender + AntiEpileptic*Li , data = temp)
summary(moodMod_ridge)

# 
# Coefficients:
#                  Estimate Std. Error t value Pr(>|t|)    
# (Intercept)      41.70715    4.43790   9.398 7.49e-11 ***
# agevisit.y       -0.72023    0.08587  -8.387 1.09e-09 ***
# gender            2.24728    1.84263   1.220   0.2313    
# AntiEpileptic    -3.71253    1.94944  -1.904   0.0656 .  
# Li               -3.92943    2.74487  -1.432   0.1617    
# AntiEpileptic:Li  5.96829    3.96027   1.507   0.1413    

```
# AntiEpileptic
```{r}
moodMod_photon <- lm(bpad_photon ~ agevisit.y + gender + AntiEpileptic , data = temp)
summary(moodMod_photon)

moodMod_brainageR <- lm(bpad_BrainAgeR~  gender + AntiEpileptic, data = temp)
summary(moodMod_brainageR)

moodMod_cnn <- lm(bpad_cnn~ agevisit.y + gender + AntiEpileptic , data = temp)
summary(moodMod_cnn)

moodMod_ridge<- lm(bpad_ridge~ agevisit.y + gender + AntiEpileptic , data = temp)
summary(moodMod_ridge)
```

# Lithium

```{r}
moodMod_photon <- lm(bpad_photon ~ agevisit.y + gender + Li , data = temp)
summary(moodMod_photon)

moodMod_brainageR <- lm(bpad_BrainAgeR~  gender + Li, data = temp)
summary(moodMod_brainageR)

moodMod_cnn <- lm(bpad_cnn~ agevisit.y + gender + Li , data = temp)
summary(moodMod_cnn)

moodMod_ridge<- lm(bpad_ridge~ agevisit.y + gender + Li , data = temp)
summary(moodMod_ridge)
```

# bpad_photon ~ mood stabilizer and emmeans
```{r}
temp$moodStabilizer <- factor(temp$moodStabilizer, levels = c("None", "AED", 'AED + Li',  "Li"))
moodMod_photon <- lm(bpad_photon ~ agevisit.y + gender + moodStabilizer , data = temp)
summary(moodMod_photon)

emm.s<-emmeans(moodMod_photon, trt.vs.ctrl ~   moodStabilizer)
counts <- temp %>% 
  filter(!is.na(moodStabilizer)) %>%
  group_by(moodStabilizer) %>% 
  summarize(count = n())
 p <- plot(emm.s,  comparisons = TRUE, , arrow.size = 5) +   
  theme_classic() +
  theme(
    axis.title.x = element_text(size = 20), 
    plot.title = element_text(size = 20, face = "bold"), 
    axis.title.y = element_text(size = 20), 
    axis.text.x = element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5, color = "black"), 
    axis.text.y = element_text(size = 18)
  )  + scale_y_discrete(labels = function(x) paste(x, "\nn=", counts$count[match(x, levels(factor(counts$moodStabilizer)))], sep = "")) + scale_x_continuous("Brain-PAD (years)")+ coord_flip()  + ylab("Medication")
 
emm_photon_df <- as.data.frame(emm.s$contrasts) %>%
  mutate(model = "PHOTON",
    lower.CL = estimate - qt(0.975, df) * SE,
    upper.CL = estimate + qt(0.975, df) * SE)
emm_photon_df


# Coefficients:
#                        Estimate Std. Error t value Pr(>|t|)    
# (Intercept)            34.85758    3.57363   9.754 5.55e-15 ***
# agevisit.y             -0.62257    0.07049  -8.833 3.10e-13 ***
# gender                  0.26758    1.58262   0.169    0.866    
# moodStabilizerAED      -8.83726    1.61974  -5.456 6.02e-07 ***
# moodStabilizerAED + Li -1.20787    2.69916  -0.447    0.656    
# moodStabilizerLi       -9.88847    2.32964  -4.245 6.20e-05 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 6.193 on 75 degrees of freedom
#   (138 observations deleted due to missingness)
# Multiple R-squared:  0.6409,	Adjusted R-squared:  0.617 
# F-statistic: 26.77 on 5 and 75 DF,  p-value: 1.984e-15

```

# bpad_ridge ~ mood stabilizerand emmeans
```{r}
temp$moodStabilizer <- factor(temp$moodStabilizer, levels = c("None", "AED", 'AED + Li',  "Li"))

moodMod_bpad_ridge<- lm(bpad_ridge ~ agevisit.y + gender + moodStabilizer , data = temp)
summary(moodMod_bpad_ridge)

emm.s<-emmeans(moodMod_bpad_ridge, trt.vs.ctrl ~   moodStabilizer)
counts <- temp %>% 
  filter(!is.na(moodStabilizer)) %>%
  group_by(moodStabilizer) %>% 
  summarize(count = n())
 p <- plot(emm.s,  comparisons = TRUE, , arrow.size = 5) +   
  theme_classic() +
  theme(
    axis.title.x = element_text(size = 20), 
    plot.title = element_text(size = 20, face = "bold"), 
    axis.title.y = element_text(size = 20), 
    axis.text.x = element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5, color = "black"), 
    axis.text.y = element_text(size = 18)
  )  + scale_y_discrete(labels = function(x) paste(x, "\nn=", counts$count[match(x, levels(factor(counts$moodStabilizer)))], sep = "")) + scale_x_continuous("Brain-PAD (years)")+ coord_flip()  + ylab("Medication")
 
emm_bpad_ridge_df <- as.data.frame(emm.s$contrasts) %>%
  mutate(model = "Multimodal",
    lower.CL = estimate - qt(0.975, df) * SE,
    upper.CL = estimate + qt(0.975, df) * SE)


# Coefficients:
#                        Estimate Std. Error t value Pr(>|t|)    
# (Intercept)            42.80244    2.89387  14.791   <2e-16 ***
# agevisit.y             -0.74666    0.05708 -13.081   <2e-16 ***
# gender                  2.10276    1.28158   1.641   0.1050    
# moodStabilizerAED      -3.15964    1.31164  -2.409   0.0185 *  
# moodStabilizerAED + Li -1.90685    2.18573  -0.872   0.3858    
# moodStabilizerLi       -3.87277    1.88650  -2.053   0.0436 *  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 5.015 on 75 degrees of freedom
#   (138 observations deleted due to missingness)
# Multiple R-squared:  0.7303,	Adjusted R-squared:  0.7123 
# F-statistic: 40.62 on 5 and 75 DF,  p-value: < 2.2e-16
```

# bpad_cnn ~ mood stabilizer and emmeans
```{r}
moodMod_bpad_cnn<- lm(bpad_cnn ~ agevisit.y + gender + moodStabilizer , data = temp)
summary(moodMod_bpad_cnn)

emm.s<-emmeans(moodMod_bpad_cnn, trt.vs.ctrl ~   moodStabilizer)
counts <- temp %>% 
  filter(!is.na(moodStabilizer)) %>%
  group_by(moodStabilizer) %>% 
  summarize(count = n())
 p <- plot(emm.s,  comparisons = TRUE, , arrow.size = 5) +   
  theme_classic() +
  theme(
    axis.title.x = element_text(size = 20), 
    plot.title = element_text(size = 20, face = "bold"), 
    axis.title.y = element_text(size = 20), 
    axis.text.x = element_text(size = 18, angle = 90, hjust = 1, vjust = 0.5, color = "black"), 
    axis.text.y = element_text(size = 18)
  )  + scale_y_discrete(labels = function(x) paste(x, "\nn=", counts$count[match(x, levels(factor(counts$moodStabilizer)))], sep = "")) + scale_x_continuous("Brain-PAD (years)")+ coord_flip()  + ylab("Medication")
  
 emm_cnn_df <- as.data.frame(emm.s$contrasts) %>%
  mutate(model = "DenseNet",
    lower.CL = estimate - qt(0.975, df) * SE,
    upper.CL = estimate + qt(0.975, df) * SE)
  
# Coefficients:
#                        Estimate Std. Error t value Pr(>|t|)    
# (Intercept)             4.39921    2.54357   1.730  0.08783 .  
# agevisit.y             -0.16204    0.05017  -3.230  0.00184 ** 
# gender                  1.32263    1.12644   1.174  0.24404    
# moodStabilizerAED      -6.70868    1.15287  -5.819 1.37e-07 ***
# moodStabilizerAED + Li -5.05162    1.92115  -2.629  0.01037 *  
# moodStabilizerLi        1.39666    1.65814   0.842  0.40230    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 4.408 on 75 degrees of freedom
#   (138 observations deleted due to missingness)
# Multiple R-squared:  0.4926,	Adjusted R-squared:  0.4588 
# F-statistic: 14.56 on 5 and 75 DF,  p-value: 5.814e-1
#  
```

# bpad_BrainAgeR ~ mood stabilizer and emmeans
```{r}
moodMod_bpad_BrainAgeR<- lm(bpad_BrainAgeR ~  gender + moodStabilizer , data = temp)
summary(moodMod_bpad_BrainAgeR)

emm.s<-emmeans(moodMod_bpad_BrainAgeR, trt.vs.ctrl ~   moodStabilizer)
counts <- temp %>% 
  filter(!is.na(moodStabilizer)) %>%
  group_by(moodStabilizer) %>% 
  summarize(count = n()) 
# plot without arrows


p <- plot(emm.s, comparisons = FALSE) +   # <-- set comparisons = FALSE
  geom_point(color = "black", size = 3.5) +  # <-- force black points
  theme_minimal() +                         # <-- switch to minimal theme
  theme(
    axis.title.x = element_text(size = 20), 
    axis.title.y = element_text(size = 20), 
    plot.title = element_text(size = 20, face = "bold"), 
    axis.text.x = element_text(size = 18, angle = 0, hjust = 1, vjust = 0.5, color = "black"), 
    axis.text.y = element_text(size = 18, color = "black"),
    legend.title = element_blank()
  ) +
  scale_y_discrete(labels = function(x) paste(x, "\nn=", counts$count[match(x, levels(factor(counts$moodStabilizer)))], sep = "")) +
  scale_x_continuous(name = "Brain-PAD (years)") +
  coord_flip() +
  ylab("Medication")
 
emm_brainageR_df <- as.data.frame(emm.s$contrasts) %>%
  mutate(model = "BrainAgeR" ,
    lower.CL = estimate - qt(0.975, df) * SE,
    upper.CL = estimate + qt(0.975, df) * SE)

aov(moodMod_bpad_BrainAgeR)

#                        Estimate Std. Error t value Pr(>|t|)  
# (Intercept)             4.88293    4.16928   1.171   0.2452  
# agevisit.y             -0.06452    0.08223  -0.785   0.4352  
# gender                  3.16031    1.84641   1.712   0.0911 .
# moodStabilizerAED      -4.20289    1.88972  -2.224   0.0292 *
# moodStabilizerAED + Li -0.84407    3.14905  -0.268   0.7894  
# moodStabilizerLi        0.99908    2.71794   0.368   0.7142  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 7.225 on 75 degrees of freedom
#   (138 observations deleted due to missingness)
# Multiple R-squared:  0.1647,	Adjusted R-squared:  0.109 
# F-statistic: 2.957 on 5 and 75 DF,  p-value: 0.01721
```

### Plot of Estimated Marginal Means of Brain-PAD across algorithms
```{r}
emm_all <- bind_rows(emm_photon_df, emm_cnn_df, emm_brainageR_df, emm_bpad_ridge_df)
ggplot(emm_all, aes(x = estimate, y = contrast
, color = model)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey") +
  geom_point(position = position_dodge(width = 0.7), size = 3) +
  geom_errorbar(aes(xmin = lower.CL, xmax = upper.CL), 
                width = 0.2, position = position_dodge(width = 0.7)) +
  labs(
    x = "Estimated Brain-PAD (years)",
    y = "Medication",
    color = "Algorithm"
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )
 #ggsave("output/figures/pairwise_comparison_meds_across_models.jpeg",dpi = 500, width = 10, height = 5)
#dev.off()

```

